---
title: "ACE"
format: html
---

```{r}
library(tidyverse)
library(readxl)
reserve <- "ACE"
```


```{r}
path_start <- here::here("submitted_data", "data", reserve)
problem_path_out <- here::here("wrangled_data", "combined_with_issues")
path_out <- here::here("wrangled_data", "CDMO")

# create the output folder if it doesn't already exist
if(!dir.exists(path_out)){
    dir.create(path_out)
}
```

## Cover  

```{r}
# dir(path_start)
# excel_sheets(here::here(path_start, "ACE Basin biomonitoring data_11Aug2021.xlsx"))
pi_cov <- read_xlsx(here::here(path_start, 
                               "ACE Basin biomonitoring data_11Aug2021_smu.xlsx"),
                    sheet = "PerCov_raw")

# because ACE uses point-intercept with 50 points,
# need to multiply raw data by 2 to get pct cover
cov_summ <- pi_cov %>% 
    group_by(Date, PlotID) %>% 
    summarize(across(c(Sediment:Mussel, Bare),
              function(x) sum(x)*2)
    ) %>% 
    select(Date, PlotID, SpaAlt, SalVir, JunRoe, Bare, everything())
```

## Height  

```{r}
hts <- read_xlsx(here::here(path_start, 
                               "ACE Basin biomonitoring data_11Aug2021_smu.xlsx"),
                    sheet = "SpHt_raw")

# ignoring SubQuad to make a new plant num column, then pivot wider
hts_wide <- hts %>% 
    group_by(Date, PlotID) %>% 
    mutate(PlantNum2 = 1:n()) %>%  
    rename(SpaAlt = "SpaAlt(cm)") %>% 
    select(Date, PlotID, PlantNum2, SpaAlt) %>% 
    pivot_wider(names_from = PlantNum2,
                names_prefix = "Spartina alterniflora_",
                values_from = SpaAlt)
```

## for data request  

### Plots  

Assigning site names based on different locations in metadata - transects 1-3 are more southern, 4-6 are more northern.  

"The biomonitoring study area, a component of the ACE Sentinel Site, is located at Edisto Island State Park, Edisto Island, SC (see map below). The study area is in the salt marshes along Big Bay Creek. Transects 1 thru 3 are in a marsh that is mainly characterized by smooth cordgrass (Spartina alterniflora) and glasswort (Salicornia virginica); while transects 4 thru 6 are in a marsh that is dominated by smooth cordgrass (Spartina alterniflora), with black needlerush (Juncus roemerianus) near the marsh-upland border."

```{r}
sites_etc <- pi_cov %>% 
    select(PlotID) %>% 
    filter(!is.na(PlotID)) %>% 
    distinct() %>% 
    separate(PlotID, into = c("TransectID", "PlotID"),
             remove = TRUE) %>% 
    mutate(SiteID = case_when(TransectID %in% c("T1", "T2", "T3") ~ "EI South",
                              TRUE ~ "EI North")) %>% 
    select(SiteID, TransectID, PlotID) %>% 
    arrange(TransectID, SiteID, PlotID)

write.csv(sites_etc, here::here("wrangled_data",
                                "combined_with_issues",
                                "ACE_plots.csv"))
```

# Cover sheet  

Changes made:  

+  insert species names rather than codes  
+  separate Transect and Plot  
+  assign SiteID based on transect  
+  separate date into ymd  

Still need to:  
  
+  insert all ID columns  

```{r}
covr <- cov_summ %>% 
    rename("Spartina alterniflora" = SpaAlt,
           "Salicornia virginica" = SalVir,
           "Juncus roemerianus" = JunRoe) %>% 
    separate(PlotID, into = c("TransectID", "PlotID"),
             sep = "-") %>% 
    mutate(SiteID = case_when(TransectID %in% c("T1", "T2", "T3") ~ "EI South",
                              TRUE ~ "EI North"),
           Year = lubridate::year(Date),
           Month = lubridate::month(Date),
           Day = lubridate::mday(Date)) %>% 
    select(Date, Year, Month, Day,
           SiteID, TransectID, PlotID,
           everything())
```


# Height sheet  

Do same things as to cover sheet. Also remove rows with no height measurements.  

```{r}
hts_wide <- hts_wide %>% 
    separate(PlotID, into = c("TransectID", "PlotID"),
             sep = "-") %>% 
    mutate(SiteID = case_when(TransectID %in% c("T1", "T2", "T3") ~ "EI South",
                              TRUE ~ "EI North"),
           Year = lubridate::year(Date),
           Month = lubridate::month(Date),
           Day = lubridate::mday(Date)) %>% 
    select(Date, Year, Month, Day,
           SiteID, TransectID, PlotID,
           everything()) %>% 
    janitor::remove_empty("cols") %>% 
    filter(!is.na(`Spartina alterniflora_1`))
```

Average heights for cover sheet  

```{r}
hts_avg <- hts_wide %>% 
    rowwise() %>% 
    mutate("Average Canopy Height_Spartina alterniflora" = round(mean(c_across(`Spartina alterniflora_1`:`Spartina alterniflora_12`), 
                                                                      na.rm = TRUE), 
                                                                 2)) %>% 
    ungroup()

hts_to_join <- hts_avg %>% 
    select(Year, Month, Day,
           SiteID, TransectID, PlotID,
           `Average Canopy Height_Spartina alterniflora`)
```


# Density sheet  

will need FlDen_raw worksheet  

```{r}
dens <- read_xlsx(here::here(path_start, 
                               "ACE Basin biomonitoring data_11Aug2021_smu.xlsx"),
                    sheet = "FlDen_raw")

# ignoring SubQuad to make a new Rep column; pivot longer to set up for pivoting wider
dens_wide <- dens %>% 
    rename(Date = SampDate...1,
           "Spartina alterniflora" = SpaAlt,
           "Salicornia virginica" = SalVir,
           "Juncus roemerianus" = JunRoe) %>% 
    group_by(Date, PlotID) %>% 
    mutate(Rep = 1:n()) %>% 
    pivot_longer(`Spartina alterniflora`:OtherVeg,
                 names_to = "Species",
                 values_to = "Value") %>% 
    select(Date, PlotID, Rep, Species, Value) %>% 
    pivot_wider(names_from = c(Species, Rep),
                names_sep = "_",
                values_from = Value) %>% 
    select(Date, PlotID,
           starts_with("Spartina"),
           starts_with("Salicornia"),
           starts_with("Juncus"),
           starts_with("Other")) %>% 
    separate(PlotID, into = c("TransectID", "PlotID"),
             sep = "-") %>% 
    mutate(SiteID = case_when(TransectID %in% c("T1", "T2", "T3") ~ "EI South",
                              TRUE ~ "EI North"),
           Year = lubridate::year(Date),
           Month = lubridate::month(Date),
           Day = lubridate::mday(Date)) %>% 
    select(Date, Year, Month, Day,
           SiteID, TransectID, PlotID,
           everything())
```


## Averaging for joining to cover sheet  

```{r}
dens_to_join <- dens_wide %>% 
    pivot_longer(-c(Date:PlotID),
                 names_to = c("Species", "Rep"),
                 names_sep = "_",
                 values_to = "Density") %>% 
    filter(!is.na(Density)) %>% 
    group_by(Year, Month, Day, 
             SiteID, TransectID, PlotID,
             Species) %>% 
    summarize(mean = round(mean(Density), 2)) %>% 
    pivot_wider(names_from = Species,
                names_prefix = "Density_",
                values_from = mean) %>% 
    select(Year:PlotID,
           "Density_Spartina alterniflora",
           "Density_Salicornia virginica",
           "Density_Juncus roemerianus",
           everything())

```


# Cover sheet - joins  



## Create F_ columns for Cover sheet  

Don't actually add them yet because we want them behind the density and height columns, but these are the names.

```{r}
spps_in_covr <- names(covr)[8:ncol(covr)]
f_names <- paste("F", spps_in_covr, sep = "_")
```

```{r}
covr_sheet <- left_join(covr, dens_to_join) %>% 
    left_join(hts_to_join) %>% 
    ungroup() %>% 
    select(-Date)

# add F_ columns
covr_sheet[ , f_names] <- NA

```

Need to add in additional ID columns and Notes, but cover, density, and height columns are all here.  


```{r}
# check for dupes
janitor::get_dupes(covr_sheet, Year, Month, Day, SiteID, TransectID, PlotID)
```


Dens and Hts sheets both have fewer rows than cover sheet; make sure to get all the same rows in all sheets  
