---
title: "Individual Reserve Template"
author: "Kim Cressman"
date: "2022-05-25; latest update `r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_depth: 2
        code_folding: hide
---

Use this template to set up the wrangling for an individual reserve's files.  


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

```{r}
library(tidyverse)
library(readxl)
library(writexl)
library(leaflet)
library(rgdal)
library(stringi)  # for stri_sub_replace()
```

# Setup  

## Helpful functions  

Because I'll be doing this repetitively on similar data files....  

```{r}
source(here::here("R", "sourced", "00_helper_functions.R"))
```

## Data path  

Update for each reserve: all start at  `here::here("submitted_data", "data")` but it varies by reserve.   

```{r}
path_start <- here::here("submitted_data", "data", "APA", "Data")
path_out <- here::here("wrangled_data", "APA")
```

## Read and combine data  

What are the files?  

```{r}
dir(path_start)  
```

Which ones to read in?  **MODIFY FOR EACH RESERVE**  

```{r}
to_read <- dir(path_start)[c(1:6, 8)]
```

Read into a list and combine  **MODIFY FOR DIFFERENT DATA TYPES**  

CDMO format:  

```{r}
dat_in <- list()

for(i in seq_along(to_read)){
    dat_in[[i]] <- read_cdmo(here::here(path_start, to_read[i])) 
}

dat_all <- bind_rows(dat_in)
```

# CDMO formatted input  

## Check station/plot names  

```{r}
unique(dat_all$Reserve)
dat_all %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct() %>% 
    knitr::kable()
```


Fix any issues here.  

```{r}
dat_all <- dat_all %>% 
    mutate(Reserve = case_when(Reserve != "APA" ~ "APA",
                               TRUE ~ Reserve))
```


## Column wrangling  

Deal with columns that were named different things in different files; also deal with date formats.  

```{r}
names(dat_all)
```


+  typo: `Averave Canopy Height` values will be transferred to `Average Canopy Height`.  
+  `Percent Cover` is a column name in 2018 only. Corresponds to `Cover` in other years.  
+  `Orthometric Height`, `Height`, and `Height Relative to MLLW` exist as columns and are all empty. `Height` is used specifically for mangroves, and the other two are retained (with underscores) in the Namaste template.  

```{r}
dat_all <- dat_all %>% 
    mutate(`Average Canopy Height` = case_when(
        !is.na(`Averave Canopy Height`) ~ `Averave Canopy Height`,
        TRUE ~ `Average Canopy Height`
    ),
    Cover = case_when(!is.na(`Percent Cover`) ~ `Percent Cover`,
                      TRUE ~ Cover),
    Year = lubridate::year(Date),
    Month = lubridate::month(Date),
    Day = lubridate::mday(Date),
    Date = format(Date, "%m/%d/%Y")
    ) %>% 
    select(-`Averave Canopy Height`, -`Percent Cover`)
```

```{r}
names(dat_all)
```

## Mangroves or SAV?  

Mangrove species are: *Avicennia germinans*, *Laguncularia racemosa*, *Rhizophora mangle*. Check for these before discarding a `Height` or `Diameter` column.   

The following sums will be 0 if none of the genus names were found in the species column. This is only set to detect the first few letteres in case of misspellings, so if any sums are above 0, it deserves further investigation.  

```{r}
spp_all <- sort(unique(dat_all$Species))
knitr::kable(spp_all)
sum(str_detect("Avic", spp_all))
sum(str_detect("Lagunc", spp_all))
sum(str_detect("Rhizo", spp_all))

```

Also check for anything other than 'E' in 'Type'.

```{r}
unique(dat_all$Type)
```



### Output prep  

## CDMO format (long)  

Select and put column names in CDMO order  

```{r}
dat_cdmo <- dat_all %>% 
    select(
        "Reserve",
        "Type",
        "Date",
        "SiteID",
        "TransectID",
        "PlotID",
        "Subplot",
        "Rep",
        "SSAM-1" = SentinelSite,
        "Lat",
        "Long",
        "Distance",
        "Orthometric Height",
        "Height Relative to MLLW",
        "Species",
        "Cover",
        "Density",
        "Maximum Canopy Height",
        "Average Canopy Height",
        "Diameter",
        "Height",
        "QAQC",
        "Notes" = Extra_1
    )
```


## Quality Checks  

```{r}
dat_cdmo %>% 
    select(Reserve:Species, QAQC) %>%
    janitor::get_dupes() %>% 
    View()

# suspect the 2nd instance of this should have been PlotID = 7 and this was a typo  
dat_cdmo %>% 
    filter(SiteID == "PC",
           TransectID == 3,
           PlotID == 5,
           Species == "Symphyotrichum tenuifolium") %>% 
    View()


# actual different values entered for this one: cover is 0.1 vs 20;
# average canopy height NA vs. 1.65 
dat_cdmo %>% 
    filter(SiteID == "LSM",
           TransectID == 1,
           PlotID == 4,
           Species == "Spartina cynosuroides") %>% 
    View()
```

```{r}
stn_dupes <- dat_cdmo %>% 
    select(Reserve, SiteID, TransectID, PlotID,
           Distance) %>% 
    distinct() %>% 
    janitor::get_dupes(-Distance)

dat_cdmo %>% 
    select(Reserve, SiteID, TransectID, PlotID,
           Distance) %>% 
    distinct() %>% 
    filter(SiteID == "PC", TransectID == 2)
# Pretty sure the Distance of 60 is the typo. How often does it show up?  

dat_cdmo %>% 
    select(SiteID, TransectID, PlotID,
           Distance) %>% 
    filter(SiteID == "PC", TransectID == 2) %>% 
    group_by(SiteID, TransectID, PlotID,
           Distance) %>% 
    tally()

# only shows up once; okay will change that to 80
# where is it though
dat_cdmo %>% 
    filter(SiteID == "PC", TransectID == 2,
           PlotID == 5, Distance == 60) %>% 
    View()

# 4/23/2018, Juncus roemerianus
```

## Correct discrepancies.  

1.  04/07/2015, PC, Transect 3, Plot 5, Symphyotrichum tenuifolium. Rows 412 and 423 in the [uncorrected] APA_CDMO.xlsx spreadsheet; I believe row 423 should be plot 7 rather than 5. This is row 422 in the current data frame. Hardcoding it is giving me heartburn but it's the best I can do right now.  

Lat and Long were also copied in for Plot 5 so are corrected here as well.  

29.59603
-85.02567

```{r}
dat_cdmo$PlotID[422] <- 7
dat_cdmo$Lat[422] <- 29.59603
dat_cdmo$Long[422] <- -85.02567
```


2.  10/03/2016, LSM, Transect 1, Plot 4 has two rows for Spartina cynosuroides. I don’t think this is a plot typo, but cover and density are different between the two rows – so maybe a species typo? It’s rows 907 and 908 in the [uncorrected] APA_CDMO spreadsheet: rows 906 and 907 in the data frame.   

```{r}

```


3.  Distance of 60 at PC-2-5 on 4/23/2018. Row 1617 in data frame.    

```{r}
dat_cdmo$Distance[1617] <- 80
```



## Write the CDMO data frame to a file.  

```{r}
write_xlsx(dat_cdmo, path = here::here(path_out, "APA_CDMO.xlsx"),
           format_headers = TRUE)
```


# NaMaSTe tables  

## Set up the station table  

Get rid of any NAs in lat/long; we want station characteristics.  

```{r}
station_table <- dat_cdmo %>% 
    select(Reserve,
           SiteID,
           TransectID,
           PlotID,
           Lat,
           Long,
           Type,
           "SSAM-1") %>% 
    filter(!is.na(Lat),
           !is.na(Long)) %>% 
    distinct()
```

Do any plots have multiple rows?  

This code led me to find incorrect Lat/Long assoicated with an incorrect plot ID; corrected up above so this code should no longer turn up problems.  

```{r}
station_table %>% 
    group_by(Reserve, SiteID, TransectID, PlotID) %>% 
    tally() %>% 
    arrange(desc(n)) %>% 
    knitr::kable()

# PC-3-7 had multiple  
station_table %>% 
    group_by(Reserve, SiteID, TransectID, PlotID) %>% 
    filter(SiteID == "PC", TransectID == 3, PlotID == 7) %>% 
    View()

dat_cdmo %>% 
    select(Reserve, SiteID, TransectID, PlotID, Lat, Long) %>% 
    filter(SiteID == "PC", TransectID == 3, PlotID == 7) %>% 
    group_by(Reserve, SiteID, TransectID, PlotID, Lat, Long) %>% 
    tally() %>% 
    View()
```

Uncorrected version (has been updated):  

Slightly different coordinates for Lat/Long at PC-3-7.  

Only one instance of these coordinates: 29.59617
-85.02579

And 85 of these: 29.59603
-85.02567

I suspect a typo.  

```{r}
dat_cdmo %>% 
    select(Reserve, SiteID, TransectID, PlotID, Date, Lat, Long) %>% 
    filter(Lat == 29.59617) %>% 
    View()
```

Yes, those are coordinates for Plot 5. And it's 4/7/2015, when the plotID typo existed. Need to change those coordinates too then. (Note: updated above.)   





Are all stations actually represented? If the below output is FALSE, there's a problem to fix. If TRUE, proceed.     

```{r}
stns_all <- dat_cdmo %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct()

nrow(stns_all) == nrow(station_table)
```


## Cover worksheet  

```{r}
cvr <- dat_cdmo %>% 
    select(Reserve:Cover) %>% 
    pivot_wider(names_from = Species,
                values_from = Cover)

cvr_qaqc <- dat_cdmo %>% 
    select(Reserve:Species, QAQC) %>%
    pivot_wider(names_from = Species,
                names_prefix = "F_",
                values_from = QAQC)



```


```{r}
# mutate(Date = lubridate::mdy(Date),
#            Year = lubridate::year(Date),
#            Month = lubridate::month(Date),
#            Day = lubridate::mday(Date)) %>% 
#     select(-Date) %>% 
```



## Supplement the station table  

Pull out the earliest Orthometric Height, Height at MLLW, Distance to Water, and Distance along transect, if they exist.  

For APA, there is a `Distance` column, which corresponds to the NaMaSTe `Distance_along_Transect` column.  




