---
title: "APA - data that has been through CDMO QAQC"
author: "Kim Cressman"
date: "2022-05-25; latest update `r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_depth: 2
        code_folding: hide
---

Use this template to set up the wrangling for an individual reserve's files.  

```{r}
library(tidyverse)
library(readxl)
library(writexl)
library(leaflet)
library(rgdal)
library(stringi)  # for stri_sub_replace()
```

# Setup  

## Helpful functions  

Because I'll be doing this repetitively on similar data files....  

```{r}
source(here::here("R", "sourced", "00_helper_functions.R"))
```

## Data path  

Update for each reserve: all start at  `here::here("submitted_data", "data")` but it varies by reserve.   

Also create an output folder for each reserve in the `wrangled data` folder and update the `path_out` below.  

```{r}
path_start <- here::here("submitted_data", "data", "APA")
problem_path_out <- here::here("wrangled_data", "combined_with_issues")
path_out <- here::here("wrangled_data", "APA")

# create the output folder if it doesn't already exist
if(!dir.exists(path_out)){
    dir.create(path_out)
}
```


What are the files?  

```{r}
dir(path_start)  
```

Which ones to read in?  **MODIFY FOR EACH RESERVE**  

```{r}
files_all <- dir(path_start) 

# figure out which ones match the pattern APAVEG[yyyy].xlsx
# \\d for digit; {4} for 'exactly 4 times'; \\. for . only (otherwise it's a class)
# stack overflow regex tutorial: https://stackoverflow.com/questions/4736/learning-regular-expressions
files_matching <- str_which(files_all, "APAVEG\\d{4}\\.xlsx")

to_read <- files_all[files_matching]
```

Read into a list and combine  **MODIFY FOR DIFFERENT DATA TYPES**  

# CDMO formatted input  

This section will compile code to read in and combine files that are already in CDMO format. Then a long CDMO-format file will be generated, after reconciling column names, dates, and putting things in order. The NaMaSTe template file will be generated from the CDMO-format file.  

```{r}
dat_in <- list()

for(i in seq_along(to_read)){
    dat_in[[i]] <- read_cdmo(here::here(path_start, to_read[i])) 
}

dat_all <- bind_rows(dat_in)
```



# Checks  

## Column names and types?  

Look for different column names for the same thing.  

```{r}
names(dat_all)
```

Still have:  

+  Averave Canopy Height - transfer to Average Canopy Height  
+  Percent Cover; transfer to Cover  
+  Newly noticed: SentinelSite should be transferred to SSAM-1.  

No extra columns with notes past the QAQC column.  

**all fixed below**  


## Duplicates in date-site-transect-plot-species  

Include subplot if used by reserve. Want an empty tibble below.    

```{r}
# subplot included
dat_all %>% 
   group_by(Date, SiteID, TransectID, PlotID, Subplot, Species) %>% 
    tally() %>% 
    filter(n > 1) %>% 
    select(Date:Species, n)

# # no subplot included
# dat_all %>% 
#    group_by(Date, SiteID, TransectID, PlotID,  Species) %>% 
#     tally() %>% 
#     filter(n > 1) %>% 
#     select(Date:Species, n)
```

Still have Spartina cynosuroides and Unvegetated dupes. No longer have mistaken plot ID.  

## Check station/plot names  

```{r}
unique(dat_all$Reserve)
dat_all %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct() %>% 
    knitr::kable()
```

Still have 'Un' as a Reserve code somewhere.    

### 'Distance' discrepancies  

```{r}
stn_dupes <- dat_all %>% 
    select(Reserve, SiteID, TransectID, PlotID,
           Distance) %>% 
    distinct() %>% 
    janitor::get_dupes(-Distance)

stn_dupes
```

PC 2-5 has distances of 60 and 80.  
PC 3-7 has distances of 60 and 40.  

```{r}
# Distances for Transect 2
dat_all %>% 
    select(SiteID, TransectID, PlotID,
           Distance) %>% 
    filter(SiteID == "PC", TransectID == 2) %>% 
    group_by(SiteID, TransectID, PlotID,
           Distance) %>% 
    tally()

# 60 only shows up once; should be 80


# # where is it though
# dat_all %>% 
#     filter(SiteID == "PC", TransectID == 2,
#            PlotID == 5, Distance == 60) %>% 
#     View()

# 4/23/2018, Juncus roemerianus



# Distances for Transect 3
dat_all %>% 
    select(SiteID, TransectID, PlotID,
           Distance) %>% 
    filter(SiteID == "PC", TransectID == 3) %>% 
    group_by(SiteID, TransectID, PlotID,
           Distance) %>% 
    tally()

# 40 only shows up once; should be 60. I bet this is from when it was logged as plot 5 instead of plot 7.  

# find it
# dat_all %>%
#     filter(SiteID == "PC", TransectID == 3,
#            PlotID == 7, Distance == 40) %>%
#     View()

# 4/7/2015, Symphyotrichum tenuifolium
```


### Lat/Long discrepancies  

```{r}
stn_dupes2 <- dat_all %>% 
    select(Reserve, SiteID, TransectID, PlotID,
           Lat, Long) %>% 
    distinct() %>% 
    janitor::get_dupes(-Lat, -Long)

stn_dupes2

# find the incorrect ones
dat_all %>% 
    select(SiteID, TransectID, PlotID,
           Lat, Long) %>% 
    filter(SiteID == "PC", TransectID == 3, PlotID == 7) %>% 
    group_by(SiteID, TransectID, PlotID,
           Lat, Long) %>% 
    tally()
```

One incorrect Lat/Long for PC-3-7, probably left over from when that plot was mistakenly entered as 5. Correct coords are:  

29.59603
-85.02567



## Mangroves or SAV present?  

Mangrove species are: *Avicennia germinans*, *Laguncularia racemosa*, *Rhizophora mangle*. Check for these before discarding a `Height` or `Diameter` column.   

The following sums will be 0 if none of the genus names were found in the species column. This is only set to detect the first few letteres in case of misspellings, so if any sums are above 0, it deserves further investigation.  

```{r}
spp_all <- sort(unique(dat_all$Species))
knitr::kable(spp_all)
sum(str_detect("Avic", spp_all))
sum(str_detect("Lagunc", spp_all))
sum(str_detect("Rhizo", spp_all))

```

Also check for anything other than 'E' in 'Type'.

```{r}
unique(dat_all$Type)
```


# Fix discrepancies  

```{r}
dat_all <- dat_all %>% 
    mutate(Reserve = case_when(Reserve != "APA" ~ "APA",
                               TRUE ~ Reserve),
           `Average Canopy Height` = case_when(
               !is.na(`Averave Canopy Height`) ~ `Averave Canopy Height`,
               TRUE ~ `Average Canopy Height`),
           Cover = case_when(!is.na(`Percent Cover`) ~ `Percent Cover`,
                             TRUE ~ Cover),
           `SSAM-1` = case_when(!is.na(SentinelSite) ~ SentinelSite,
                                TRUE ~ `SSAM-1`)
           ) %>% 
    select(-`Averave Canopy Height`, -`Percent Cover`, -SentinelSite)
```






# Format  

## Column wrangling  


If there are any necessary changes, make them here. Also pull out columns for Month, Day, and Year; and format Date as mm/dd/yyyy for the CDMO output file.  

```{r}
dat_all <- dat_all %>% 
    mutate(
        Year = lubridate::year(Date),
        Month = lubridate::month(Date),
        Day = lubridate::mday(Date),
        Date = format(Date, "%m/%d/%Y")
    ) 
```

```{r}
names(dat_all)
```




## CDMO format (long)  

CDMO column names, in order. If columns had different names, or there were additional ones, account for that here.    

```{r}
dat_cdmo <- dat_all %>% 
    select(
        "Reserve",
        "Type",
        "Date",
        "Year",
        "Month",
        "Day",
        "SiteID",
        "TransectID",
        "PlotID",
        "Subplot",
        "Rep",
        "SSAM-1",
        "Lat",
        "Long",
        "Distance",
        "Orthometric Height",
        "Height Relative to MLLW",
        "Species",
        "Cover",
        "Density",
        "Maximum Canopy Height",
        "Average Canopy Height",
        "Diameter",
        "Height",
        "QAQC",
        everything()
    )
```



## Write the CDMO data frame to a file. **Modify for each reserve**  

```{r}
write.csv(dat_cdmo, here::here(problem_path_out, "APA_issues.csv"), 
          row.names= FALSE,
          na = "")
```


```{r}
write_xlsx(dat_cdmo, path = here::here(path_out, "APA_CDMO.xlsx"),
           format_headers = TRUE)
```


# NaMaSTe tables  

Get rid of any NAs in lat/long; we want station characteristics.  

```{r}
station_table <- dat_cdmo %>% 
    select(Reserve,
           SiteID,
           TransectID,
           PlotID,
           Lat,
           Long,
           Type,
           "SSAM-1") %>% 
    filter(!is.na(Lat),
           !is.na(Long)) %>% 
    distinct()
```

Do any plots have multiple rows?  

```{r}
station_table %>% 
    group_by(Reserve, SiteID, TransectID, PlotID) %>% 
    tally() %>% 
    arrange(desc(n)) %>% 
    knitr::kable()
```

Are all stations actually represented? If the below output is FALSE, there's a problem to fix. If TRUE, proceed.     

```{r}
stns_all <- dat_cdmo %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct()

nrow(stns_all) == nrow(station_table)
```









# OLD     




### Output prep  

## CDMO format (long)  

Select and put column names in CDMO order  

```{r}
dat_cdmo <- dat_all %>% 
    select(
        "Reserve",
        "Type",
        "Date",
        "SiteID",
        "TransectID",
        "PlotID",
        "Subplot",
        "Rep",
        "SSAM-1" = SentinelSite,
        "Lat",
        "Long",
        "Distance",
        "Orthometric Height",
        "Height Relative to MLLW",
        "Species",
        "Cover",
        "Density",
        "Maximum Canopy Height",
        "Average Canopy Height",
        "Diameter",
        "Height",
        "QAQC",
        "Notes" = Extra_1
    )
```





## Correct discrepancies.  

1.  04/07/2015, PC, Transect 3, Plot 5, Symphyotrichum tenuifolium. Rows 412 and 423 in the [uncorrected] APA_CDMO.xlsx spreadsheet; I believe row 423 should be plot 7 rather than 5. This is row 422 in the current data frame. Hardcoding it is giving me heartburn but it's the best I can do right now.  

Lat and Long were also copied in for Plot 5 so are corrected here as well.  

29.59603
-85.02567

```{r}
dat_cdmo$PlotID[422] <- 7
dat_cdmo$Lat[422] <- 29.59603
dat_cdmo$Long[422] <- -85.02567
```



3.  Distance of 60 at PC-2-5 on 4/23/2018. Row 1617 in data frame.    

```{r}
dat_cdmo$Distance[1617] <- 80
```


2.  2nd issue to come up but moved below the others because I'm going to remove a row.  

10/03/2016, LSM, Transect 1, Plot 4 has two rows for Spartina cynosuroides. I don’t think this is a plot typo, but cover and density are different between the two rows – so maybe a species typo? It’s rows 907 and 908 in the [uncorrected] APA_CDMO spreadsheet: rows 906 and 907 in the data frame.   

--per emails with the reserve, veg monitoring is conducted by contractors and the reserve recommends flagging these as suspect. (5/26/22)  

--unfortunately to pivot, I need to only have one row with all these identifiers and species. Because row 906 (the first entry) only shows cover of 0.1 and no density, heights, or other data; and removing row 907 which has cover of 20% would greatly change the total percent cover for the plot; I believe I'll flag row 907 as suspect, add a note into the 'notes' column, and then remove row 906.  

```{r}
dat_cdmo$QAQC[907] <- "<1>"
dat_cdmo$Notes[907] <- "Originally two rows for this Spartina cynosuroides in this plot on this date. First row recorded 0.1 for cover and no other information; that row was removed and the row containing 20 for cover and 1.65 for canopy height was retained."
dat_cdmo <- dat_cdmo[-906, ]
```



4. More dupes when attempting to pivot:  

```{r}
more_dupen <- dat_cdmo %>% 
    select(Reserve:Cover) %>% 
    pivot_wider(names_from = Species,
                values_from = Cover,
                values_fn = length)

summary(more_dupen)

# seems to be unvegetated

more_dupen %>% 
    filter(Unvegetated == 2) %>% 
    View()

more_dupen %>% 
    filter(Unvegetated == 2) %>% 
    select(Reserve:Distance) %>% 
    left_join(dat_cdmo) %>% 
    select(Date:PlotID, Species, Cover) %>% 
    filter(Species == "Unvegetated") %>% 
    View()
```

4 instances where 'Unvegetated' was entered twice for a plot on a date. With the double entries, total cover adds up to ~100%, so rather than removing information, will add cover for both rows together, insert <1> QAQC flag, and detail in 'Notes'. These instances are:  10/04/2016 PC-2-7;   04/18/2017 PC-2-2;   04/18/2017 PC 2-7;   04/23/2018 PC 3-1.

# FIX BEFORE PROCEEDING  


## Write the CDMO data frame to a file.  

```{r}
write_xlsx(dat_cdmo, path = here::here(path_out, "APA_CDMO.xlsx"),
           format_headers = TRUE)
```


# NaMaSTe tables  

## Set up the station table  

Get rid of any NAs in lat/long; we want station characteristics.  

```{r}
station_table <- dat_cdmo %>% 
    select(Reserve,
           SiteID,
           TransectID,
           PlotID,
           Lat,
           Long,
           Type,
           "SSAM-1") %>% 
    filter(!is.na(Lat),
           !is.na(Long)) %>% 
    distinct()
```

Do any plots have multiple rows?  

This code led me to find incorrect Lat/Long assoicated with an incorrect plot ID; corrected up above so this code should no longer turn up problems.  

```{r}
station_table %>% 
    group_by(Reserve, SiteID, TransectID, PlotID) %>% 
    tally() %>% 
    arrange(desc(n)) %>% 
    knitr::kable()

# # PC-3-7 had multiple  
# station_table %>% 
#     group_by(Reserve, SiteID, TransectID, PlotID) %>% 
#     filter(SiteID == "PC", TransectID == 3, PlotID == 7) %>% 
#     View()
# 
# dat_cdmo %>% 
#     select(Reserve, SiteID, TransectID, PlotID, Lat, Long) %>% 
#     filter(SiteID == "PC", TransectID == 3, PlotID == 7) %>% 
#     group_by(Reserve, SiteID, TransectID, PlotID, Lat, Long) %>% 
#     tally() %>% 
#     View()
```

Uncorrected version (has been updated):  

Slightly different coordinates for Lat/Long at PC-3-7.  

Only one instance of these coordinates: 29.59617
-85.02579

And 85 of these: 29.59603
-85.02567

I suspect a typo.  

```{r}
# dat_cdmo %>% 
#     select(Reserve, SiteID, TransectID, PlotID, Date, Lat, Long) %>% 
#     filter(Lat == 29.59617) %>% 
#     View()
```

Yes, those are coordinates for Plot 5. And it's 4/7/2015, when the plotID typo existed. Need to change those coordinates too then. (Note: updated above.)   





Are all stations actually represented? If the below output is FALSE, there's a problem to fix. If TRUE, proceed.     

```{r}
stns_all <- dat_cdmo %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct()

nrow(stns_all) == nrow(station_table)
```


## Cover worksheet  

```{r}
cvr <- dat_cdmo %>% 
    select(Reserve:Cover) %>% 
    pivot_wider(names_from = Species,
                values_from = Cover,
                values_fn = length)

cvr_qaqc <- dat_cdmo %>% 
    select(Reserve:Species, QAQC) %>%
    pivot_wider(names_from = Species,
                names_prefix = "F_",
                values_from = QAQC)



```


```{r}
# mutate(Date = lubridate::mdy(Date),
#            Year = lubridate::year(Date),
#            Month = lubridate::month(Date),
#            Day = lubridate::mday(Date)) %>% 
#     select(-Date) %>% 
```



## Supplement the station table  

Pull out the earliest Orthometric Height, Height at MLLW, Distance to Water, and Distance along transect, if they exist.  

For APA, there is a `Distance` column, which corresponds to the NaMaSTe `Distance_along_Transect` column.  




