---
title: "CBM-OPC"
format: html
---

```{r}
library(tidyverse)
library(readxl)
library(writexl)
```

```{r}
# starting path for data
path_start <- here::here("submitted_data", "data")
file_in <- here::here(path_start, "CBM", "Veg", "Otter Point Creek", 
                      "MASTER_OPC_EV_Transect_Data_060512_LC.xls")

# excel_sheets(file_in)
```

from mapping script - to read in file and deal with coordinates  

## Station Table  

```{r}
cbm_prep <- read_xls(file_in, sheet = "Site Descriptions",
                    skip = 9)
names(cbm_prep)[4:5] <- c("Lat", "Long")

cbm_opc <- read_xls(file_in, sheet = "Site Descriptions",
                     skip = 12,
                     col_names = FALSE)
names(cbm_opc) <- names(cbm_prep)
rm(cbm_prep)

# use "Current Transect Name" column; that's what's in the data sheets
cbm_opc <- cbm_opc %>% 
    rename(TransectID = `Current Transect Name`) %>% 
    fill(TransectID) %>% 
    mutate(Reserve = "CBM",
           SiteID = "OPC",
           Lat = round(Lat, 5),
           Long = -1 * round(Long, 5)) %>% 
    select(-"...6", -"...7") %>% 
    select(Reserve, SiteID,
           TransectID,
           PlotID = Plot,
           Lat, Long,
           everything())

stn_tbl <- cbm_opc %>% 
    mutate("SSAM-1" = ifelse(!is.na(PlotID), "yes", NA),
           Orthometric_Height = NA,
           Height_Relative_to_MLLW = NA,
           Type = ifelse(!is.na(PlotID), "E", NA)) %>% 
    select(Reserve, SiteID, TransectID, PlotID,
           Lat, Long, Type,
           Orthometric_Height,
           Height_Relative_to_MLLW,
           Distance = `Distance (m)`,
           `SSAM-1`,
           everything())
```

## Cover  

Date entry problem for WDCT4-P5 in 2012 - entered as "7/17/201". Updating to "41107" which is the Excel numeric date that corresponds.  

```{r}
covr_in <- read_excel(file_in, sheet = "Cover Data", skip = 12)
covr_in <- covr_in %>% 
    mutate(Date = ifelse(Date == "07/17/201", "41107", Date),
           Date = janitor::excel_numeric_to_date(as.numeric(Date))) 
```

## NA date issues - find rows where there are numbers  

Typically an NA for date means the plot couldn't be found, so in theory any numbers entered are typos. There are some, however, so need to pull these out and investigate them.  

```{r}
# date is NA. keep only columns that have an entry *somewhere*.
NAdates <- covr_in %>% 
    filter(is.na(Date)) %>% 
    janitor::remove_empty("cols")

# index the rows that don't have any entries for the species and remove.
# do rowSums on "not NAs" in the species columns and 
# keep those which are greater than 0
inds <- which(rowSums(!is.na(NAdates[5:ncol(NAdates)])) > 0)
    
to_investigate <- NAdates[inds, ]

```


## Density  

```{r}
dens_in <- read_excel(file_in, sheet = "Stem Density", skip = 11)
dens <- dens_in %>% 
    janitor::remove_empty("cols")
```

```{r}
# add Density_ prefix for joining in cover sheet
dens_to_join <- dens
densNames <- paste("Density", names(dens), sep = "_")
names(dens_to_join)[6:ncol(dens_to_join)] <- densNames[6:ncol(dens_to_join)]
```


## Heights  

```{r}
hts_in <- read_excel(file_in, sheet = "Max Height Data", skip = 9)
hts <- hts_in %>% 
    janitor::remove_empty("cols")
```


```{r}
# add Maximum Canopy Height_ prefix for joining in cover sheet
hts_to_join <- hts
htsNames <- paste("Maximum Canopy Height", names(hts), sep = "_")

names(hts_to_join)[6:ncol(hts_to_join)] <- htsNames[6:ncol(hts_to_join)]
```


## Species List  

```{r}
spp_covr <- read_excel(file_in, 
                       sheet = "Cover Data", 
                       skip = 11, n_max = 2,
                       col_names = FALSE)
names(spp_covr) <- names(covr_in)
spp_covr <- spp_covr %>% 
    filter(row_number() != 2) %>% 
    select(6:93)

spp_names <- spp_covr %>% 
    pivot_longer(everything(),
                 names_to = "Species in column names",
                 values_to = "Common name")
```



# Generate Cover sheet  


## Add columns of station table info  


```{r}
stn_to_join <- stn_tbl %>% 
    select(Reserve, SiteID, TransectID,
           PlotID, Orthometric_Height,
           Height_Relative_to_MLLW,
           Distance) %>% 
    filter(!is.na(PlotID))
```

```{r}
covr_sheet <- covr_in %>% 
    rename(TransectID = Transect,
           PlotID = Plot) %>% 
    left_join(stn_to_join) %>% 
    select(Reserve, Region, SiteID,
           Year, Date, 
           TransectID, PlotID,
           Orthometric_Height,
           Height_Relative_to_MLLW,
           Distance,
           Notes = Comments,
           everything())
```



## Create F_ columns for Cover sheet  







## Join Density and Height  


