---
title: "CBM-OPC"
format: html
---

```{r}
library(tidyverse)
library(readxl)
library(writexl)
```

```{r}
# starting path for data
path_start <- here::here("submitted_data", "data")
file_in <- here::here(path_start, "CBM", "Veg", "Otter Point Creek", 
                      "MASTER_OPC_EV_Transect_Data_060512_LC.xls")

# excel_sheets(file_in)
```

from mapping script - to read in file and deal with coordinates  

## Station Table  

```{r}
cbm_prep <- read_xls(file_in, sheet = "Site Descriptions",
                    skip = 9)
names(cbm_prep)[4:5] <- c("Lat", "Long")

cbm_opc <- read_xls(file_in, sheet = "Site Descriptions",
                     skip = 12,
                     col_names = FALSE)
names(cbm_opc) <- names(cbm_prep)
rm(cbm_prep)

# use "Current Transect Name" column; that's what's in the data sheets
cbm_opc <- cbm_opc %>% 
    rename(TransectID = `Current Transect Name`) %>% 
    fill(TransectID) %>% 
    mutate(Reserve = "CBM",
           SiteID = "OPC",
           Lat = round(Lat, 5),
           Long = -1 * round(Long, 5)) %>% 
    select(-"...6", -"...7") %>% 
    select(Reserve, SiteID,
           TransectID,
           PlotID = Plot,
           Lat, Long,
           everything())

stn_tbl <- cbm_opc %>% 
    mutate("SSAM-1" = ifelse(!is.na(PlotID), "yes", NA),
           Orthometric_Height = NA,
           Height_Relative_to_MLLW = NA,
           Type = ifelse(!is.na(PlotID), "E", NA)) %>% 
    fill(Heading, .direction = "down") %>% 
    filter(!is.na(PlotID)) %>% 
    select(-Transect) %>% 
    select(Reserve, SiteID, TransectID, PlotID,
           Lat, Long, Type,
           Orthometric_Height,
           Height_Relative_to_MLLW,
           Distance = `Distance (m)`,
           `SSAM-1`,
           everything())
```

## Cover  

Date entry problem for WDCT4-P5 in 2012 - entered as "7/17/201". Updating to "41107" which is the Excel numeric date that corresponds.  

```{r}
covr_in <- read_excel(file_in, sheet = "Cover Data", skip = 12)
covr_in <- covr_in %>% 
    mutate(Date = ifelse(Date == "7/17/201", "41107", Date),
           Date = ifelse(Date == "07/17/201", "41107", Date),
           Date = janitor::excel_numeric_to_date(as.numeric(Date)),
           Year = ifelse(is.na(Year), 2008, Year)) 
```

## NA date issues   

### get rid of 0s  

Per Kyle, these were typos (there were only 2). Originally found with the following code. Just muscling the 0-to-NA thing.  


```{r}
# date is NA. keep only columns that have an entry *somewhere*.
NAdates <- covr_in %>% 
    filter(is.na(Date)) %>% 
    select(-Comments) %>% 
    janitor::remove_empty("cols")

# index the rows that don't have any entries for the species and remove.
# do rowSums on "not NAs" in the species columns and 
# keep those which are greater than 0
inds <- which(rowSums(!is.na(NAdates[5:ncol(NAdates)])) > 0)
    
to_investigate <- NAdates[inds, ]

```

...and change.  

```{r}
rows_to_change <- covr_in %>% 
    mutate(rownum = row_number()) %>% 
    filter(Transect == "HHT5",
           Plot == "P3",
           Year %in% c(2008, 2009)) %>% 
    select(rownum) %>% 
    unlist()

covr_in[rows_to_change, "Bidens aristosa/frondosa"] <- NA
```


### attach correct dates  

Some transects were sampled over multiple dates. I'm assigning the first date for each transect for simplicity (even though the actual sample date may have been one of the other days - these are plots that were not sampled).  

Also removing all of 2016 rows in this process (Date2 will be NA).  

```{r}
transectDates <- covr_in %>% 
    select(Transect, Year, Date) %>% 
    distinct() %>% 
    filter(!is.na(Date)) %>% 
    rename(Date2 = Date)

twoDates <- transectDates %>% 
    select(Transect, Year) %>% 
    duplicated()

transectSingle <- transectDates[-which(twoDates == TRUE), ]

covr_in <- left_join(covr_in, transectSingle) %>% 
    filter(!is.na(Date2)) %>% 
    mutate(Date = case_when(is.na(Date) ~ Date2,
                            TRUE ~ Date)) %>% 
    select(-Date2) %>% 
    rename("Sparganium americanum" = "Sparganium spp.")
```



## Density  

Renaming:  

+  Bidens aristosa to Bidens aristosa/frondosa in both density and height, per conversation with Kyle D. Both species are present and impossible to tell apart unless flowering. We were planning to go with Bidens sp. but another species is present (Bidens laevis). Density only.    
+  Decadon verticillatus to Decodon verticillatus (original was typo). Density only.  
+  Dichanthelum spp. to Dichanthelium spp. (typo). Height only.  
+  Sparganium spp. to Sparganium americanum (per Kyle, americanum is only species present). COVER only (performed above).  
+  Unknown Grass spp. to Unknown spp. Both density and height.  

```{r}
dens_in <- read_excel(file_in, sheet = "Stem Density", skip = 11)
dens <- dens_in %>% 
    janitor::remove_empty("cols") %>% 
    rename("Bidens aristosa/frondosa" = "Bidens aristosa",
           "Decodon verticillatus" = "Decadon verticillatus"
           # "Dichanthelium spp." = "Dichanthelum spp.",
           # "Sparganium americanum" = "Sparganium spp.",
           ) %>% 
    rowwise() %>% 
    mutate(`Unknown spp.` = sum(`Unknown spp.`, `Unknown Grass spp.`, na.rm = TRUE)) %>% 
    select(-`Unknown Grass spp.`)
```

```{r}
# add Density_ prefix for joining in cover sheet
dens_to_join <- dens
densNames <- paste("Density", names(dens), sep = "_")
names(dens_to_join)[6:ncol(dens_to_join)] <- densNames[6:ncol(dens_to_join)]
```


## Heights  

```{r}
hts_in <- read_excel(file_in, sheet = "Max Height Data", skip = 9)
hts <- hts_in %>% 
    janitor::remove_empty("cols") %>% 
    rename(# "Bidens aristosa/frondosa" = "Bidens aristosa",
           # "Decodon verticillatus" = "Decadon verticillatus"
           "Dichanthelium spp." = "Dichanthelum spp."
           # "Sparganium americanum" = "Sparganium spp."
           ) %>% 
    rowwise() %>% 
    mutate(`Unknown spp. 2` = sum(`Unknown spp.`, `Unknown Grass spp.`, na.rm = TRUE)) %>% 
    select(-`Unknown Grass spp.`, -`Unknown spp.`) %>% 
    rename("Unknown spp." = "Unknown spp. 2")
```


```{r}
# add Maximum Canopy Height_ prefix for joining in cover sheet
hts_to_join <- hts
htsNames <- paste("Maximum Canopy Height", names(hts), sep = "_")

names(hts_to_join)[6:ncol(hts_to_join)] <- htsNames[6:ncol(hts_to_join)]
```


## Species List  

```{r}
spp_covr <- read_excel(file_in, 
                       sheet = "Cover Data", 
                       skip = 11, n_max = 2,
                       col_names = FALSE)
names(spp_covr) <- names(covr_in)
spp_covr <- spp_covr %>% 
    filter(row_number() != 2) %>% 
    select(6:93)

spp_names <- spp_covr %>% 
    pivot_longer(everything(),
                 names_to = "Species",
                 values_to = "Common name")
```



# Generate Cover sheet  


## Add columns of station table info  


```{r}
stn_to_join <- stn_tbl %>% 
    select(Reserve, SiteID, TransectID,
           PlotID, Orthometric_Height,
           Height_Relative_to_MLLW,
           Distance) %>% 
    filter(!is.na(PlotID))
```

```{r}
covr_sheet <- covr_in %>% 
    rename(TransectID = Transect,
           PlotID = Plot) %>% 
    left_join(stn_to_join) %>% 
    select(Reserve, Region, SiteID,
           Year, Date, 
           TransectID, PlotID,
           Orthometric_Height,
           Height_Relative_to_MLLW,
           Distance,
           Notes = Comments,
           everything())
```



## Create F_ columns for Cover sheet  

Don't actually add them yet because we want them behind the density and height columns, but these are the names.

```{r}
spps_in_covr <- names(covr_sheet)[12:ncol(covr_sheet)]
f_names <- paste("F", spps_in_covr, sep = "_")

# test <- covr_sheet
# test[ , f_names] <- NA
```



## Join Density and Height  

Remember to get rid of NA dates (first check to make sure there's not real data in there)  

```{r}
sum(is.na(dens_to_join$Date))
dens_to_join %>% 
    filter(is.na(Date)) %>% 
    janitor::remove_empty("cols") 

sum(is.na(hts_to_join$Date))
hts_to_join %>% 
    filter(is.na(Date)) %>% 
    janitor::remove_empty("cols") # %>% View()
```

Everything with NA for date in the density and heights sheets also has no data, so can be disregarded.  

```{r}
dens_to_join <- dens_to_join %>% 
    filter(!is.na(Date)) %>% 
    select(-Density_Comments) %>% 
    rename(TransectID = Transect,
           PlotID = Plot)

hts_to_join <- hts_to_join %>% 
    filter(!is.na(Date)) %>% 
    select(-`Maximum Canopy Height_Comments`) %>% 
    rename(TransectID = Transect,
           PlotID = Plot)
```

# Cover sheet for real  

Then remove the date column and make Year, Month, Day ones.  

```{r}
covr_sheet <- left_join(covr_sheet, dens_to_join) %>% 
    left_join(hts_to_join)

# add F_ columns
covr_sheet[ , f_names] <- NA

covr_sheet <- covr_sheet %>% 
    mutate(Year = lubridate::year(Date),
           Month = lubridate::month(Date),
           Day = lubridate::mday(Date)) %>% 
    relocate(c(Month, Day), .after = Year) %>% 
    select(-Date)
    
```


# Density sheet  

Want to keep all the same rows, but only the density columns  

Then remove the 'Density' prefix  

```{r}
dens_sheet <- covr_sheet %>% 
    select(Reserve:Notes,
           starts_with("Density_"))

names(dens_sheet) <- str_remove(names(dens_sheet), "Density_")
```



# Heights sheet  

Same for heights  

```{r}
hts_sheet <- covr_sheet %>% 
    select(Reserve:Notes,
           starts_with("Maximum Canopy Height_"))

names(hts_sheet) <- str_remove(names(hts_sheet), "Maximum Canopy Height_")
```


# Compile workbook  

```{r}
out_path <- here::here("wrangled_data", "NMST", "CBM-OPC_NMST.xlsx")
write_xlsx(list(Station_Table = stn_tbl,
                Species_Names = spp_names,
                Cover = covr_sheet,
                Density = dens_sheet,
                Height = hts_sheet),
           path = out_path,
           format_headers = TRUE)
```

