---
title: "ELK"
output: html_document
date: "2023-02-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)
library(writexl)
library(leaflet)
library(rgdal)
library(stringi)  # for stri_sub_replace()

reserve <- "ELK"
```

# Setup  

## Helpful functions  

Because I'll be doing this repetitively on similar data files....  

```{r}
source(here::here("R", "sourced", "00_helper_functions.R"))
```

## Data path  

Just one file to read here.  



```{r}
# path_start <- here::here("submitted_data", "data", reserve, 
#                          "CBNERRVA Veg Data GI Reserve for Chris Peter.xlsx")

path_start <- here::here("submitted_data", "data", reserve)

problem_path_out <- here::here("wrangled_data", "combined_with_issues")

path_out <- here::here("wrangled_data", "CDMO")

# create the output folder if it doesn't already exist
if(!dir.exists(path_out)){
    dir.create(path_out)
}
```


Read into a list and combine  **MODIFY FOR DIFFERENT DATA TYPES**  

# CDMO formatted input  

This section will compile code to read in and combine files that are already in CDMO format. Then a long CDMO-format file will be generated, after reconciling column names, dates, and putting things in order. The NaMaSTe template file will be generated from the CDMO-format file.  

```{r}
dat <- read_xlsx(here::here(path_start, "Transects_Compiled.xlsx"), 
                 sheet = "quads compiled")
```

Per Kerstin in 'Elk spp' file (in response to Jan/Feb 2023 data request): "we separate succulent and woody because former is more responsive to short-term changes, but you don't need to bother - delete juicy/woody and just use the overall one"   

Looks like lat and long are switched, for 2016, 2018, and part of 2021. May have to do with certain sites?? Okay, yes, looks like Hester is the only site with *correct* lat/long (e.g. lat around 36, not -121)

```{r}
dat <- dat %>% 
    select(-Juicy_Pickleweed,
           -Woody_Pickleweed,
           -High_Marsh,
           -Category,
           -(Marsh_Cover:Burrow_Density),  # keep Pickleweed_Height
           -(Pickleweed_Biomass_Index:Seaward_Pickleweed_Elevation))
```


## Fix lat/longs  

```{r}
dat <- dat %>% 
    mutate(Lat = case_when(Site == "Hester" ~ Latitude,
                           .default = Longitude),
           Long = case_when(Site == "Hester" ~ Longitude,
                            .default = Latitude)) %>% 
    relocate(Lat, .after = Latitude) %>% 
    relocate(Long, .after = Longitude) %>% 
    select(-Latitude, -Longitude)
```



## Latin names  

```{r}
spps <- read_xlsx(here::here(path_start, "ELK spp.xlsx"),
                  col_names = c("Species_from", "Species_to", "native", "comments")) %>% 
    mutate(Species_to = case_when(is.na(Species_to) ~ Species_from,
                                  .default = Species_to))

dat_names <- tibble("current" = names(dat)) %>% 
    left_join(spps, by = c("current" = "Species_from")) %>% 
    mutate(new = case_when(!is.na(Species_to) ~ Species_to,
                           .default = current)) 

names(dat) <- dat_names$new 
```


## Additional renaming and massaging

```{r}
dat_all <- dat %>% 
    rename(SiteID = Site,
           TransectID = Transect,
           PlotID = Quadrat,
           Orthometric_Height = Elevation,
           "Maximum Canopy Height_Salicornia pacifica" = Pickleweed_Height,
           Distance_along_transect = Transect_Distance) %>% 
    mutate(Reserve = "ELK",
           Year = lubridate::year(Date),
           Month = lubridate::month(Date),
           Day = lubridate::mday(Date),
           TransectID = case_when(is.na(TransectID) ~ 1,
                                  .default = TransectID)) %>% 
    select(-Date) %>% 
    select(Reserve, 
           Year, Month, Day,
           SiteID, TransectID, PlotID,
           everything())
```



# Checks  

## Duplicates in date-site-transect-plot-species  

Want an empty tibble.  

```{r}
janitor::get_dupes(dat_all, Year, Month, Day, SiteID, TransectID, PlotID)
```




## Check station/plot names  

```{r}
unique(dat_all$Reserve)
```

```{r}
dat_all %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct() %>% 
    knitr::kable()
```


### 'Distance' discrepancies  

```{r}
stn_dupes <- dat_all %>% 
    select(Reserve, SiteID, TransectID, PlotID,
           Distance = Distance_along_transect) %>% 
    distinct() %>% 
    janitor::get_dupes(-Distance)

stn_dupes
```




## Column wrangling  



# NaMaSTe format  


```{r}
# function to pick first non-NA reading for any given plot ("at establishment")
establ <- function(df, x){
    
    x_vec <- df %>% 
        select({{x}})
    
    if(sum(!is.na(x_vec)) == 0){
        out <- df %>% 
            select(SiteID, TransectID, PlotID, {{x}}) %>% 
            distinct() 
    } else {
        out <- df %>% 
            select(Date, SiteID, TransectID, PlotID, {{x}}) %>% 
            distinct() %>% 
            group_by(SiteID, TransectID, PlotID) %>% 
            filter(!is.na({{x}})) %>%
            filter(Date == min(Date)) %>% 
            select(-Date) 
    }

    out
}
```

Renaming possibilities:  

```{r}
col_matching <- c("Orthometric_Height" = "Orthometric Height",
                  "Height_Relative_to_MLLW" = "Height Relative to MLLW",
                  "Ht" = "Canopy Height",
                  "Ht" = "Average Canopy Height",
                  "Ht" = "Maximum Canopy Height")

# deal with different types of canopy height measurement
ht_type <- "Maximum Canopy Height"
```



## Identifiers  

Should be the same between all sheets  

```{r}
id_names <- c("Reserve", "SiteID", 
             "Year", "Month", "Day",
             "TransectID", "PlotID",
             "Orthometric_Height", "Height_Relative_to_MLLW",
             "Distance_along_transect", "Notes")

# only use ID columns that exist in the data frame
id_names <- id_names[which(id_names %in% names(dat_all))]
```


# Station Table  

Start. Doesn't have elevations or some of the required columns.

```{r}
stn_tbl <- dat_all %>% 
    select(Reserve, SiteID, TransectID,
           PlotID, Lat, Long, Distance_along_transect) %>% 
    distinct()
```


# Cover sheet  


## Cover data  

```{r}
covr <- dat_all %>% 
    select(all_of(id_names), 
           Bare:`Maximum Canopy Height_Salicornia pacifica`)

```

### F_ columns  

```{r}
F_cols <- paste("F", names(select(covr, Bare:Chenopodiaceae)), sep = "_")

covr[ , F_cols] <- NA
```



## Density data  

None; making a blank worksheet.

```{r}
dens <- dat_all %>% 
    select(all_of(id_names)) 

# apply(dens, MARGIN = 2, FUN = function(x) sum(!is.na(x))) 
# no empty species columns 
```


## Height data  

```{r}
hts <- dat_all %>% 
    select(all_of(id_names), 
           "Salicornia pacifica" = "Maximum Canopy Height_Salicornia pacifica") 
```


## Join them all  

```{r}
covr_sheet <- full_join(covr_wide, dens) %>% 
    full_join(hts) %>% 
    select(all_of(id_names), everything()) %>% 
    relocate(Notes, .after = Distance) %>% 
    relocate(starts_with("F_"), .after = last_col())

```

## If NA columns, remove  

This can happen when reserves don't sample a site on a certain date, but they have a placeholder row in the data where species is blank. During pivoting, an 'NA' column is created. So is an 'F_NA' column.  

```{r}
if("NA" %in% names(covr_sheet)){
    covr_sheet$"NA" <- NULL
}

if("F_NA" %in% names(covr_sheet)){
    covr_sheet$"F_NA" <- NULL
}
```


Check again for dupes  

```{r}
janitor::get_dupes(covr_sheet, Year, Month, Day,
                   SiteID, TransectID, PlotID)
```



# Station Table  

Everything in this table is the first recorded value.  

# DEFINITELY NEED TO MODIFY THIS    


```{r}
# there must be a better way to do this, but it works.

stns <- dat %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct() %>% 
    arrange(SiteID, TransectID, PlotID)
lats <- establ(dat, Lat) %>% 
    group_by(SiteID, TransectID, PlotID) %>% 
    summarize(Lat = mean(Lat))   # p13A had 2 values for the first date
longs <- establ(dat, Long) %>% 
    group_by(SiteID, TransectID, PlotID) %>% 
    summarize(Long = mean(Long))   # p13A had 2 values for the first date
ssam1 <- establ(dat, `SSAM-1`)
dists <- establ(dat, Distance)
orth_hts <- establ(dat, Orthometric_Height) 
mllw_hts <- establ(dat, Height_Relative_to_MLLW)
types <- establ(dat, Type)

stn_tbl <- full_join(stns, lats) %>% 
    full_join(longs) %>% 
    full_join(types) %>% 
    full_join(orth_hts) %>% 
    full_join(mllw_hts) %>%
    full_join(dists) %>% 
    full_join(ssam1)

```



# Species List  

```{r}
spp_names <- dat %>% 
    select(Species) %>% 
    distinct() %>% 
    arrange(Species)
```




# Write to Excel  

# Compile workbook  

```{r}
out_path <- here::here("wrangled_data", "NMST", "ELK_NMST.xlsx")
write_xlsx(list(Station_Table = stn_tbl,
                Species_Names = spp_names,
                Cover = covr,
                Density = dens,
                Height = hts),
           path = out_path,
           format_headers = TRUE)
```