---
title: "GND unveg"
output: html_document
date: "2023-03-08; updated 2023-11-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

GND doesn't have explicitly identified "unvegetated" covers. Adding them here, as 100 - (sum of identified covers). Seems easiest to do this from the Namaste formatted file. Will then add to the CDMO file.  

11-29-23 update: starting from latest CDMO file, which has the site name issue resolved (spal/spalt and clamj/clmaj). In starting manual NMST formatting, staff noticed certain rows didn't add up to 100 - which is mostly because the calculations didn't get performed entirely correctly the first time (???). We also found two plots where the total was >100, and those will be fixed in this script as well.  

```{r}
library(tidyverse)
library(readxl)
library(writexl)
```


```{r}
dat <- read_xlsx(here::here("wrangled_data", 
                            "CDMO",
                            "GND_CDMO_not100s.xlsx"),
                 na = c("", "NA")) %>% 
    filter(!is.na(Year))
```


```{r}
# unvegetated is in here
"Unvegetated" %in% unique(dat$Species)

# but sites are right
unique(dat$SiteID)

dat %>%  
    summarize(.by = c(Year, Month, Day, SiteID, TransectID, PlotID),
              total = sum(Cover, na.rm = TRUE))  %>%  
    arrange(desc(total))  %>%  
    View()
```


## Fix the plots > 100  

Per JP in emails 11/29-11/30/2023: 2017 Juro Mid should be 5% Juro, 10% Sppatens, and 0 for everything else. 2017 clmaj, change "unknown" to 2.5.    

```{r}
# juro mid site, change 60 and 55 to 0
rowind <- dat %>% 
    mutate(rownum = row_number()) %>% 
    filter(Year == 2017,
        SiteID == "juro mid",
        TransectID == 1,
        PlotID == 1,
        Species %in% c("Panicum virgatum", "Cladium jamaicense")) %>% 
    select(rownum) %>% 
    unlist()
dat[rowind,
     "Cover"] <- 0

# change unknown to 2.5
rowind <- dat %>% 
    mutate(rownum = row_number()) %>% 
    filter(Year == 2017,
        SiteID == "clmaj",
        TransectID == 3,
        PlotID == 2,
        Species == "Unknown") %>% 
    select(rownum) %>% 
    unlist()
dat[rowind,
     "Cover"] <- 2.5


# check it
dat %>%  
    summarize(.by = c(Year, Month, Day, SiteID, TransectID, PlotID),
              total = sum(Cover, na.rm = TRUE))  %>%  
    arrange(desc(total))  %>%  
    View()
# good, nothing left over 100
```

## Find plots < 100  

Generate data for unvegetateds....  

```{r}
Totals <- dat %>%  
    summarize(.by = c(Year, Month, Day, SiteID, TransectID, PlotID),
              total = sum(Cover, na.rm = TRUE)) 

Total_live <- dat %>%  
    filter(Species != "Unvegetated") %>% 
    summarize(.by = c(Year, Month, Day, SiteID, TransectID, PlotID),
              total_live = sum(Cover, na.rm = TRUE))

All <- left_join(dat, Totals) %>% 
    left_join(Total_live)

All2 <- All %>% 
    mutate(Cover2 = case_when(Species == "Unvegetated" & total != 100 ~ 100 - total_live,
                              .default = Cover))

dat2 <- All2 %>% 
    mutate(Cover = Cover2) %>% 
    select(-total, -total_live, -Cover2)
```

```{r}
# check it
dat2 %>%  
    summarize(.by = c(Year, Month, Day, SiteID, TransectID, PlotID),
              total = sum(Cover, na.rm = TRUE))  %>%  
    arrange(total)  %>%  
    View()

dat <- dat2
```


Re-write CDMO file:  

```{r}
write_xlsx(dat, here::here("wrangled_data",
                           "CDMO",
                           "GND_CDMO.xlsx"),
           format_headers = TRUE)
```

