---
title: "GTM"
author: "Kim Cressman"
date: "2022-05-25; latest update `r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_depth: 2
        code_folding: hide
---

Use this template to set up the wrangling for an individual reserve's files.  


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

```{r}
library(tidyverse)
library(readxl)
library(writexl)
library(leaflet)
library(rgdal)
library(stringi)  # for stri_sub_replace()

reserve <- "GTM"
```

# Setup  

## Helpful functions  

Because I'll be doing this repetitively on similar data files....  

```{r}
source(here::here("R", "sourced", "00_helper_functions.R"))
```

## Data path  

Update for each reserve: all start at  `here::here("submitted_data", "data")` but it varies by reserve.   

Also create an output folder for each reserve in the `wrangled data` folder and update the `path_out` below.  

```{r}
path_start <- here::here("submitted_data", "data", reserve)
problem_path_out <- here::here("wrangled_data", "combined_with_issues")
path_out <- here::here("wrangled_data", "CDMO")

# create the output folder if it doesn't already exist
if(!dir.exists(path_out)){
    dir.create(path_out)
}
```


What are the files?  

```{r}
dir(path_start)  
```

Which ones to read in?  **MODIFY FOR EACH RESERVE**  

```{r}
files_all <- dir(path_start) 

# figure out which ones match the pattern APAVEG[yyyy].xlsx:
# str_which(files_all, "APAVEG\\d{4}\\.xlsx")
# \\d for digit; {4} for 'exactly 4 times'; \\. for . only (otherwise it's a class)
# stack overflow regex tutorial: https://stackoverflow.com/questions/4736/learning-regular-expressions
files_matching <- str_which(files_all, "GTMVEG\\d{4}\\.xlsx")

to_read <- files_all[files_matching]
```

Read into a list and combine  **MODIFY FOR DIFFERENT DATA TYPES**  

# CDMO formatted input  

This section will compile code to read in and combine files that are already in CDMO format. Then a long CDMO-format file will be generated, after reconciling column names, dates, and putting things in order. The NaMaSTe template file will be generated from the CDMO-format file.  

```{r}
dat_in <- list()


for(i in seq_along(to_read)){
    # deal with 2015 worksheet issue
    if(to_read[i] == "GTMVEG2015.xlsx"){
        dat_in_tmp <- read_cdmo(here::here(path_start, to_read[i]),
                                 worksheet = "GTM2015")
    } else {
        dat_in_tmp <- read_cdmo(here::here(path_start, to_read[i]),
                                 worksheet = NULL)
    }
     
    
    # deal with PlotID character vs. numeric issue
    if(is.character(dat_in_tmp$PlotID)){
        dat_in_tmp <- dat_in_tmp %>% 
            mutate(PlotID = as.numeric(PlotID))
    }
    
    # assign to list
    dat_in[[i]] <- dat_in_tmp
}

# 4th file filed: 
# GTMVEG2015. Has extra worksheet first: Sheet 1. Sheet to read in is 2nd, GTM2015.
# fixed with if-else statement above; if file to read is the problematic one, read that sheet. otherwise, sheet = NULL.


# # bind_rows error on PlotID
# for(i in 1:length(dat_in)){
#     cat(to_read[i], "\n", class(dat_in[[i]]$PlotID), "\n \n")
# }
# # character in all but 2020: entered as 01, 02, etc; in 2020 entered as 1, 2, etc. and are numeric
# # fixed by making character strings numeric for GTM


dat_all <- bind_rows(dat_in)
```



# Checks  

## Column names and types?  

Look for different column names for the same thing.  

```{r}
names(dat_all)
```

## Duplicates in date-site-transect-plot-species  

Include subplot if used by reserve. Want an empty tibble below.    

```{r}
# # subplot included
# dat_all %>% 
#    group_by(Date, SiteID, TransectID, PlotID, Subplot, Species) %>% 
#     tally() %>% 
#     filter(n > 1) %>% 
#     select(Date:Species, n)

# no subplot included
dat_all %>%
   group_by(Date, SiteID, TransectID, PlotID,  Species) %>%
    tally() %>%
    filter(n > 1) %>%
    select(Date:Species, n)

# 742 rows of NA for all selected columns. get rid of these.
dat_all <- dat_all %>% 
    filter(!is.na(Date))
```


## Check station/plot names  

```{r}
unique(dat_all$Reserve)
```


```{r}
dat_all %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct() %>% 
    knitr::kable()
```


### 'Distance' discrepancies  

```{r}
stn_dupes <- dat_all %>% 
    select(Reserve, SiteID, TransectID, PlotID,
           Distance) %>% 
    mutate(uniqueID = paste(SiteID, TransectID, PlotID, sep = "_")) %>% 
    distinct() %>% 
    janitor::get_dupes(-Distance)

stn_dupes
# 
# dat_all %>% 
#     mutate(uniqueID = paste(SiteID, TransectID, PlotID, sep = "_")) %>% 
#     filter(uniqueID %in% stn_dupes$uniqueID) %>% 
#     group_by(uniqueID, Distance) %>% 
#     tally()
# 
# dat_all %>% 
#     mutate(uniqueID = paste(SiteID, TransectID, PlotID, sep = "_")) %>% 
#     filter(uniqueID == "46_Lower_15") %>% 
#     View()
# 
# dat_all %>% 
#     mutate(uniqueID = paste(SiteID, TransectID, PlotID, sep = "_")) %>% 
#     filter(uniqueID == "01_C_5") %>% 
#     arrange(desc(Distance)) %>% 
#     View()
```


### Fix the above discrepancies  


```{r}
dat_all <- dat_all %>% 
    mutate(Reserve = case_when(Reserve != "GTM" ~ "GTM",
                               TRUE ~ Reserve),
           uniqueID = paste(SiteID, TransectID, PlotID, sep = "_"),
           Distance = case_when(uniqueID == "01_C_5" ~ 6,
                                uniqueID == "46_Lower_15" ~ 140,
                                TRUE ~ Distance)) %>% 
    select(-uniqueID)
```


### Lat/Long discrepancies  



## Mangroves or SAV present?  

Mangrove species are: *Avicennia germinans*, *Laguncularia racemosa*, *Rhizophora mangle*. Check for these before discarding a `Height` or `Diameter` column.   

The following sums will be 0 if none of the genus names were found in the species column. This is only set to detect the first few letteres in case of misspellings, so if any sums are above 0, it deserves further investigation.  

```{r}
spp_all <- sort(unique(dat_all$Species))
knitr::kable(spp_all)
sum(str_detect(spp_all, "Avic"))
sum(str_detect(spp_all, "Lagunc"))
sum(str_detect(spp_all, "Rhizo"))

```

Also check for anything other than 'E' in 'Type'.

```{r}
unique(dat_all$Type)
```



# Fix discrepancies  


## Column wrangling  

Pull out columns for Month, Day, and Year; and format Date as mm/dd/yyyy for the CDMO output file.  

```{r}
dat_all <- dat_all %>% 
    mutate(
        Year = lubridate::year(Date),
        Month = lubridate::month(Date),
        Day = lubridate::mday(Date),
        Date = format(Date, "%m/%d/%Y")
    ) 
```



## CDMO format (long)  

CDMO column names, in order. If columns had different names, or there were additional ones, account for that here.    

```{r}
dat_cdmo <- dat_all %>% 
    select(
        "Reserve",
        "Type",
        "Date",
        "SiteID",
        "TransectID",
        "PlotID",
        "Subplot",
        "Rep",
        "SSAM-1",
        "Lat",
        "Long",
        "Distance",
        "Orthometric Height",
        "Height Relative to MLLW",
        "Species",
        "Cover",
        "Density",
        "Canopy Height",
        "Diameter",
        "Height",
        "QAQC"
    )
```


## Write the CDMO data frame to a file. **Modify for each reserve**  

```{r}
write_xlsx(dat_cdmo, path = here::here(path_out, paste0(reserve, "_CDMO.xlsx")),
           format_headers = TRUE)
```


# NaMaSTe tables  

Get rid of any NAs in lat/long; we want station characteristics.  

```{r}
station_table <- dat_cdmo %>% 
    select(Reserve,
           SiteID,
           TransectID,
           PlotID,
           Lat,
           Long,
           Type,
           "SSAM-1") %>% 
    filter(!is.na(Lat),
           !is.na(Long)) %>% 
    distinct()
```

Do any plots have multiple rows?  

```{r}
station_table %>% 
    group_by(Reserve, SiteID, TransectID, PlotID) %>% 
    tally() %>% 
    arrange(desc(n)) %>% 
    knitr::kable()
```

Are all stations actually represented? If the below output is FALSE, there's a problem to fix. If TRUE, proceed.     

```{r}
stns_all <- dat_cdmo %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct()

nrow(stns_all) == nrow(station_table)
```