---
title: "JAC"
author: "Kim Cressman"
date: "2022-05-25; latest update `r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_depth: 2
        code_folding: hide
---

Use this template to set up the wrangling for an individual reserve's files.  


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

```{r}
library(tidyverse)
library(readxl)
library(writexl)
library(leaflet)
library(rgdal)
library(stringi)  # for stri_sub_replace()
```

# Setup  

## Helpful functions  

Because I'll be doing this repetitively on similar data files....  

```{r}
source(here::here("R", "sourced", "00_helper_functions.R"))
```

## Data path  

Update for each reserve: all start at  `here::here("submitted_data", "data")` but it varies by reserve.   

Also create an output folder for each reserve in the `wrangled data` folder and update the `path_out` below.  

```{r}
path_start <- here::here("submitted_data", "data", "JAC")
problem_path_out <- here::here("wrangled_data", "combined_with_issues")
path_out <- here::here("wrangled_data", "CDMO")

# create the output folder if it doesn't already exist
if(!dir.exists(path_out)){
    dir.create(path_out)
}
```

JAC has subfolders for each year of data. In each subfolder is a file `JAC yyyy Marsh Vegetation.xlsx` data file.

What are the subfolders?  

```{r}
folders_all <- dir(path_start)
subf_index <- which(nchar(folders_all) == 4)  
years <- folders_all[subf_index]
```

Which ones to read in?  **MODIFY FOR EACH RESERVE**  

```{r}
to_read <- character() 

for(i in seq_along(years)){
    to_read[i] <- here::here(path_start, years[i], 
                               paste0("JAC ", years[i], " Marsh Vegetation.xlsx"))
}

```

Read into a list and combine  **MODIFY FOR DIFFERENT DATA TYPES**  

# CDMO formatted input  

This section will compile code to read in and combine files that are already in CDMO format. Then a long CDMO-format file will be generated, after reconciling column names, dates, and putting things in order. The NaMaSTe template file will be generated from the CDMO-format file.  

Only want 'Density...17', because it's been converted to per m^2. True in all input files.  

2011 file has 2 columns for canopy height - first is in cm; 2nd is in m. Other files are only in m, so only want to keep the 2nd.  

```{r}
dat_in <- list()

for(i in seq_along(to_read)){
    
    dat_in_tmp <- read_cdmo(to_read[i]) 
    
    # 2 canopy height columns in 2011
    if(years[i] == "2011"){
        dat_in_tmp <- dat_in_tmp %>% 
            select(-"Canopy Height...18") %>% 
            rename("Canopy Height" = "Canopy Height...19")
    } 
    
    # different column title in 2013
    if(years[i] == "2013"){
        dat_in_tmp <- dat_in_tmp %>% 
            rename(SiteID = `SITE ID`) %>% 
            mutate(SiteID = as.character(SiteID))
    } 
    
    # first two years have transectID as character format, but can be turned into numbers
    if(is.character(dat_in_tmp$TransectID)){
        dat_in_tmp$TransectID <- as.numeric(dat_in_tmp$TransectID)
    }
    
    # all have 2 density columns and we only want the 2nd
    dat_in_tmp <- dat_in_tmp %>% 
        select(-"Density...16") %>% 
        rename(Density = "Density...17")
    
    dat_in[[i]] <- dat_in_tmp
}


dat_all <- bind_rows(dat_in)
```

Deal with extra columns - should be Notes. Extra_2 is just NAs, only need Extra_1.  

```{r}
dat_all <- dat_all %>% 
    select(-Extra_2) %>% 
    rename(Notes = Extra_1)
```


# Checks  

## Column names and types?  

Look for different column names for the same thing.  

```{r}
names(dat_all)
```

## Duplicates in date-site-transect-plot-species  

Include subplot if used by reserve. Want an empty tibble below.    

```{r}
# # subplot included
dat_all %>%
   group_by(Date, SiteID, TransectID, PlotID, Subplot, Species) %>%
    tally() %>%
    filter(n > 1) %>%
    select(Date:Species, n)

dupes <- dat_all %>% 
    mutate(rownum = row_number()) %>% 
    janitor::get_dupes(Date, SiteID, TransectID, PlotID, Subplot, Species)

# no subplot included
# dat_all %>%
#    group_by(Date, SiteID, TransectID, PlotID,  Species) %>%
#     tally() %>%
#     filter(n > 1) %>%
#     select(Date:Species, n) %>% 
#     View()
```


Per reserve on 1/24/2023, The second entry there (row 629 in the original data sheet; cover = 35, density = 768) should be Spartina patens.  

## Fix dupe  

```{r}
find_dupe <- dat_all %>% 
    mutate(rownum = row_number()) %>% 
    filter(PlotID == "1013-8-1",
           Species == "Spartina alterniflora",
           `% Cover` == 35)
dat_all[find_dupe$rownum, "Species"] <- "Spartina patens"
```


## Check station/plot names  

```{r}
unique(dat_all$Reserve)
dat_all %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct() %>% 
    knitr::kable()
```


## Change station/plot columns to be consistent with other reserves  

Per the metadata, transects 1-3 are in the "North Segment" of the reserve; 4-6 are in the "Central Segment"; and 7-9 are in the "South Segment". I will use these as site names (currently the SiteID column is populated with plot number-within-transect). Additionally, the current PlotID column is a unique ID; I will remove the unique ID (will be added in later per NMST protocols) and change the PlotID to what is currently in SiteID.  

```{r}
dat_all <- dat_all %>% 
    select(-PlotID) %>% 
    rename(PlotID = SiteID) %>% 
    mutate(SiteID = case_when(TransectID %in% 1:3 ~ "North Segment",
                             TransectID %in% 4:6 ~ "Central Segment",
                             TransectID %in% 7:9 ~ "South Segment",
                             TRUE ~ "Something is wrong"),
           PlotID = as.numeric(PlotID)) %>% 
    relocate(SiteID, .after = Date) %>% 
    relocate(PlotID, .after = TransectID)
```


### 'Distance' discrepancies  

```{r}
stn_dupes <- dat_all %>% 
    select(Reserve, SiteID, TransectID, PlotID,
           Distance) %>% 
    distinct() %>% 
    janitor::get_dupes(-Distance)

stn_dupes

counts <- dat_all %>% 
    select(Reserve, SiteID, TransectID, PlotID, Distance) %>% 
    group_by(Reserve, SiteID, TransectID, PlotID, Distance) %>% 
    tally()

left_join(stn_dupes, counts)
```

Transect 4, Plot 3 is labelled 66.9 only twice; 44.6 44 times. Update.  

Transect 4, Plot 4 vice versa - update 44.6 to 66.9.  

Transect 8, Plot 5 has one NA. Change to 89.2.

```{r}
dat_all <- dat_all %>% 
    mutate(Distance = case_when(TransectID == 4 & PlotID == 3 ~ 44.6,
                                TransectID == 4 & PlotID == 4 ~ 66.9,
                                TransectID == 8 & PlotID == 5 ~ 89.2,
                                TRUE ~ Distance))
```


### Lat/Long discrepancies  



## Mangroves or SAV present?  

Mangrove species are: *Avicennia germinans*, *Laguncularia racemosa*, *Rhizophora mangle*. Check for these before discarding a `Height` or `Diameter` column.   

The following sums will be 0 if none of the genus names were found in the species column. This is only set to detect the first few letteres in case of misspellings, so if any sums are above 0, it deserves further investigation.  

```{r}
spp_all <- sort(unique(dat_all$Species))
knitr::kable(spp_all)
sum(str_detect(spp_all, "Avic"))
sum(str_detect(spp_all, "Lagunc"))
sum(str_detect(spp_all, "Rhizo"))

```

Also check for anything other than 'E' in 'Type'.

```{r}
unique(dat_all$Type)
```

## Species list issues  

```{r}
spp_all
str_detect(spp_all, "^[Uu]nveg")
unvegs_all <- spp_all[str_detect(spp_all, "^[Uu]nveg")]
```

There's also at least one "Salicornia.spp." that should be "Salicornia spp." 

## Fix species  

```{r}
dat_all <- dat_all %>% 
    mutate(Species = case_when(Species %in% unvegs_all ~ "Unvegetated",
                               Species == "Salicornia.spp." ~ "Salicornia spp.",
                               TRUE ~ Species))
```

Check again for dupes, in case there were multiple unvegetated categories on a given date.  

```{r}
unveg_dupes <- dat_all %>% 
    mutate(rownum = row_number()) %>% 
    janitor::get_dupes(Date, SiteID, TransectID, PlotID, Species) %>% 
    relocate(rownum)
```

There is a dupe, transect 7 plot 5. 75% was salt panne and 15% was another type of unvegetated.   

Update the first to 90% cover and remove the other row.  

```{r}
dat_all[unveg_dupes$rownum[1], "% Cover"] <- 90

dat_all <- dat_all %>% 
    filter(row_number() != unveg_dupes$rownum[2])
```


# Fix discrepancies  


## Column wrangling  

Pull out columns for Month, Day, and Year; and format Date as mm/dd/yyyy for the CDMO output file.  

```{r}
dat_all <- dat_all %>% 
    mutate(
        Year = lubridate::year(Date),
        Month = lubridate::month(Date),
        Day = lubridate::mday(Date),
        Date = format(Date, "%m/%d/%Y")
    ) 
```



## CDMO format (long)  

CDMO column names, in order. If columns had different names, or there were additional ones, account for that here.    

```{r}
dat_cdmo <- dat_all %>% 
    select(
        "Reserve",
        "Type",
        "Date",
        "SiteID",
        "TransectID",
        "PlotID",
        "Subplot",
        "Rep",
        "Lat",
        "Long",
        "Distance",
        "Orthometric Height" = Elevation,
        "Species",
        "Cover" = `% Cover`,
        "Density",
        "Maximum Canopy Height" = `Canopy Height`,
        "Diameter",
        "Height",
        "Notes"
    )
```


## Write the CDMO data frame to a file. **Modify for each reserve**  

```{r}
write_xlsx(dat_cdmo, path = here::here(path_out, "JAC_CDMO.xlsx"),
           format_headers = TRUE)
```


# NaMaSTe tables  

Get rid of any NAs in lat/long; we want station characteristics.  

```{r}
station_table <- dat_cdmo %>% 
    select(Reserve,
           SiteID,
           TransectID,
           PlotID,
           Lat,
           Long,
           Type,
           "SSAM-1") %>% 
    filter(!is.na(Lat),
           !is.na(Long)) %>% 
    distinct()
```

Do any plots have multiple rows?  

```{r}
station_table %>% 
    group_by(Reserve, SiteID, TransectID, PlotID) %>% 
    tally() %>% 
    arrange(desc(n)) %>% 
    knitr::kable()
```

Are all stations actually represented? If the below output is FALSE, there's a problem to fix. If TRUE, proceed.     

```{r}
stns_all <- dat_cdmo %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct()

nrow(stns_all) == nrow(station_table)
```