---
title: "KAC"
author: "Kim Cressman"
date: "2022-05-25; latest update `r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_depth: 2
        code_folding: hide
---

Use this template to set up the wrangling for an individual reserve's files.  


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

```{r}
library(tidyverse)
library(readxl)
library(writexl)
library(leaflet)
library(rgdal)
library(stringi)  # for stri_sub_replace()
```

# Setup  

## Helpful functions  

Because I'll be doing this repetitively on similar data files....  

```{r}
source(here::here("R", "sourced", "00_helper_functions.R"))
```

## Data path  

Update for each reserve: all start at  `here::here("submitted_data", "data")` but it varies by reserve.   

Also create an output folder for each reserve in the `wrangled data` folder and update the `path_out` below.  

```{r}
path_start <- here::here("submitted_data", "data", "KAC", "Recent data")
path2_start <- here::here("submitted_data", "data", "KAC", "Data from CDMO")
problem_path_out <- here::here("wrangled_data", "combined_with_issues")
path_out <- here::here("wrangled_data", "KAC")

# create the output folder if it doesn't already exist
if(!dir.exists(path_out)){
    dir.create(path_out)
}
```


What are the files?  

```{r}
dir(path_start)  
dir(path2_start)  # subfolders for each year, 2010-2013
```

2016-2021 are in the 'Recent Data' folder. There are subfolders in 'Data from CDMO' for each of 2010, 2011, 2012, and 2013.  


Which ones to read in?  **MODIFY FOR EACH RESERVE**  

```{r}
to_read <- dir(path_start)[c(1, 3, 5, 7, 10)]

in2010 <- here::here(path2_start, "KACVEG2010", "KACVEG2010.xlsx")
in2011 <- here::here(path2_start, "KACVEG2011", "KACVEG2011.xlsx")
in2012 <- here::here(path2_start, "KACVEG2012", "KACVEG2012.xlsx")
in2013 <- here::here(path2_start, "KACVEG2013", "KACVEG2013.xlsx")

to_read2 <- c(in2010, in2011, in2012, in2013)
```

Read into a list and combine  **MODIFY FOR DIFFERENT DATA TYPES**  

# CDMO formatted input  

This section will compile code to read in and combine files that are already in CDMO format. Then a long CDMO-format file will be generated, after reconciling column names, dates, and putting things in order. The NaMaSTe template file will be generated from the CDMO-format file.  


### 2016-2021:  

```{r}
dat_in <- list()

for(i in seq_along(to_read)){
    dat_in[[i]] <- read_cdmo(here::here(path_start, to_read[i])) 
}

dat_all1 <- bind_rows(dat_in)
```


### 2010-2013  

```{r}
dat_in2 <- list()

for(i in seq_along(to_read2)){
    dat_in2[[i]] <- read_cdmo(to_read2[i]) 
}

dat_all2 <- bind_rows(dat_in2)
```


## Join all together  

```{r}
dat_all <- bind_rows(dat_all1, dat_all2)
```



# Checks  

## Column names and types?  

Look for different column names for the same thing.  

```{r}
names(dat_all)
```

To reconcile:  

+  Cover / % Cover  
+  Elevation / Orthometric Height   

Several values in the `% Cover` Column are "R": per the metadata, this means "Trace", or <1%. For our purposes, am converting to 0.1.  

```{r}
dat_all <- dat_all %>% 
    mutate(`% Cover` = case_when(`% Cover` == "R" ~ "0.1",
                                 TRUE ~ `% Cover`),
           `% Cover` = as.numeric(`% Cover`),
           Cover = case_when(!is.na(`% Cover`) ~ `% Cover`,
                             TRUE ~ Cover),
           `Orthometric Height` = case_when(!is.na(Elevation) ~ (Elevation/100),
                                     TRUE ~ `Orthometric Height`)
           ) %>% 
    select(-`% Cover`, -`Elevation`)
```



## Duplicates in date-site-transect-plot-species  

Include subplot if used by reserve. Want an empty tibble below.    

```{r}
# # subplot included
# dat_all %>% 
#    group_by(Date, SiteID, TransectID, PlotID, Subplot, Species) %>% 
#     tally() %>% 
#     filter(n > 1) %>% 
#     select(Date:Species, n)
# 
# no subplot included
dat_all %>%
   group_by(Date, SiteID, TransectID, PlotID,  Species) %>%
    tally() %>%
    filter(n > 1) %>%
    select(Date:Species, n)
```

5 instances of duplicated "unvegetated":  

+  8/17/2010 FR-T3 FR23  
+  8/18/2010 FR-T4 FR28  
+  8/18 2010 FR-T5 FR54  
+  8/18/2010 FR-T6 FR57  
+  8/7/2013 BS-T14 BS47  


## Check station/plot names  

```{r}
unique(dat_all$Reserve)
```

```{r}
dat_all %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct() %>% 
    knitr::kable()
```


Fix any issues here.  

```{r}

```


## Mangroves or SAV present?  

Mangrove species are: *Avicennia germinans*, *Laguncularia racemosa*, *Rhizophora mangle*. Check for these before discarding a `Height` or `Diameter` column.   

The following sums will be 0 if none of the genus names were found in the species column. This is only set to detect the first few letteres in case of misspellings, so if any sums are above 0, it deserves further investigation.  

```{r}
spp_all <- sort(unique(dat_all$Species))
knitr::kable(spp_all)
sum(str_detect(spp_all, "Avic"))
sum(str_detect(spp_all, "Lagunc"))
sum(str_detect(spp_all, "Rhizo"))

```

Also check for anything other than 'E' in 'Type'.

```{r}
unique(dat_all$Type)
```


## Column wrangling  



```{r}
names(dat_all)
```

If there are any necessary changes, make them here. Also pull out columns for Month, Day, and Year; and format Date as mm/dd/yyyy for the CDMO output file.  

```{r}
dat_all <- dat_all %>% 
    mutate(
        Year = lubridate::year(Date),
        Month = lubridate::month(Date),
        Day = lubridate::mday(Date),
        Date = format(Date, "%m/%d/%Y")
    ) 
```

```{r}
names(dat_all)
```




## CDMO format (long)  

CDMO column names, in order. If columns had different names, or there were additional ones, account for that here.    

```{r}
dat_cdmo <- dat_all %>% 
    select(
        "Reserve",
        "Type",
        "Date",
        "Year",
        "Month",
        "Day",
        "SiteID",
        "TransectID",
        "PlotID",
        "Subplot",
        "Rep",
        "SSAM-1",
        "Lat",
        "Long",
        "Distance",
        "Orthometric Height",
        "Height Relative to MLLW",
        "Species",
        "Cover",
        "Density",
        "Canopy Height",
        "Diameter",
        "Height",
        "QAQC"
    )
```

## Quality Checks  

```{r}
dat_cdmo %>% 
    select(Reserve:Species, QAQC) %>%
    janitor::get_dupes() %>% 
    View()
```

```{r}
dat_cdmo %>% 
    select(Reserve, SiteID, TransectID, PlotID,
           Distance) %>% 
    distinct() %>% 
    janitor::get_dupes(-Distance)
```


## Correct discrepancies  


## Write the CDMO data frame to a file. **Modify for each reserve**  

```{r}
write.csv(dat_cdmo, here::here(problem_path_out, "KAC_issues.csv"), 
          row.names= FALSE,
          na = "")
```

```{r}
write_xlsx(dat_cdmo, path = here::here(path_out, "Reserve_CDMO.xlsx"),
           format_headers = TRUE)
```


# NaMaSTe tables  

Get rid of any NAs in lat/long; we want station characteristics.  

```{r}
station_table <- dat_cdmo %>% 
    select(Reserve,
           SiteID,
           TransectID,
           PlotID,
           Lat,
           Long,
           Type,
           "SSAM-1") %>% 
    filter(!is.na(Lat),
           !is.na(Long)) %>% 
    distinct()
```

Do any plots have multiple rows?  

```{r}
station_table %>% 
    group_by(Reserve, SiteID, TransectID, PlotID) %>% 
    tally() %>% 
    arrange(desc(n)) %>% 
    knitr::kable()
```

Are all stations actually represented? If the below output is FALSE, there's a problem to fix. If TRUE, proceed.     

```{r}
stns_all <- dat_cdmo %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct()

nrow(stns_all) == nrow(station_table)
```