---
title: "NIW"
author: "Kim Cressman"
date: "2022-05-25; latest update `r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_depth: 2
        code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

```{r}
library(tidyverse)
library(readxl)
library(writexl)
library(leaflet)
library(rgdal)
library(stringi)  # for stri_sub_replace()
```

# Setup  

## Helpful functions  

Because I'll be doing this repetitively on similar data files....  

```{r}
source(here::here("R", "sourced", "00_helper_functions.R"))

# function to pick first non-NA reading for any given plot ("at establishment")
establ <- function(df, x){
    
    x_vec <- df %>% 
        select({{x}})
    
    if(sum(!is.na(x_vec)) == 0){
        out <- df %>% 
            select(SiteID, TransectID, PlotID, {{x}}) %>% 
            distinct() 
    } else {
        out <- df %>% 
            select(Date, SiteID, TransectID, PlotID, {{x}}) %>% 
            distinct() %>% 
            group_by(SiteID, TransectID, PlotID) %>% 
            filter(!is.na({{x}})) %>%
            filter(Date == min(Date)) %>% 
            select(-Date)
    }
        
    out
}
```

## Data path  

Update for each reserve: all start at  `here::here("submitted_data", "data")` but it varies by reserve.   

Also create an output folder for each reserve in the `wrangled data` folder and update the `path_out` below.  

```{r}
reserve <- "NIW"
path_start <- here::here("submitted_data", "data", "NIW")
problem_path_out <- here::here("wrangled_data", "combined_with_issues")
path_out <- here::here("wrangled_data", "CDMO")
out_path <- here::here("wrangled_data", "NMST", paste0(reserve, "_NMST.xlsx"))
```


What are the files?  

```{r}
dir(path_start)  
```

Which ones to read in?  **MODIFY FOR EACH RESERVE**  

```{r}
step1 <- str_starts(dir(path_start), "NIWVEG")
step2 <- str_detect(dir(path_start), ".xlsx")
step3 <- step1 + step2


to_read <- dir(path_start)[which(step3 == 2)]
```

Read into a list and combine  **MODIFY FOR DIFFERENT DATA TYPES**  

# CDMO formatted input  

This section will compile code to read in and combine files that are already in CDMO format. Then a long CDMO-format file will be generated, after reconciling column names, dates, and putting things in order. The NaMaSTe template file will be generated from the CDMO-format file.  

```{r}
dat_in <- list()

for(i in seq_along(to_read)){
    dat_in[[i]] <- read_cdmo(here::here(path_start, to_read[i])) 
}

dat_all <- bind_rows(dat_in) %>% 
    rename(Orthometric_Height = `Orthometric Height`,
           Height_Relative_to_MLLW = `Height Relative to MLLW`)
```


# Checks  

## Column names and types?  

Look for different column names for the same thing.  

```{r}
names(dat_all)
```

## Duplicates in date-site-transect-plot-species  

Include subplot if used by reserve. Want an empty tibble below.    

```{r}
# subplot included
dat_all %>% 
   group_by(Date, SiteID, TransectID, PlotID, Subplot, Species) %>% 
    tally() %>% 
    filter(n > 1) %>% 
    select(Date:Species, n)

# # no subplot included
# dat_all %>% 
#    group_by(Date, SiteID, TransectID, PlotID,  Species) %>% 
#     tally() %>% 
#     filter(n > 1) %>% 
#     select(Date:Species, n)
```


## Check station/plot names  

```{r}
unique(dat_all$Reserve)
dat_all %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct() %>% 
    knitr::kable()
```


Fix any issues here.  

```{r}

```


## Mangroves or SAV present?  

Mangrove species are: *Avicennia germinans*, *Laguncularia racemosa*, *Rhizophora mangle*. Check for these before discarding a `Height` or `Diameter` column.   

The following sums will be 0 if none of the genus names were found in the species column. This is only set to detect the first few letteres in case of misspellings, so if any sums are above 0, it deserves further investigation.  

```{r}
spp_all <- sort(unique(dat_all$Species))
knitr::kable(spp_all)
sum(str_detect(spp_all, "Avic"))
sum(str_detect(spp_all, "Lagunc"))
sum(str_detect(spp_all, "Rhizo"))

```

Also check for anything other than 'E' in 'Type'.

```{r}
unique(dat_all$Type)
```




## Column wrangling  

Check for columns that may have been named differently in different files.  

```{r}
names(dat_all)
```

If there are any necessary changes, make them here. Also pull out columns for Month, Day, and Year; and format Date as mm/dd/yyyy for the CDMO output file.  

```{r}
dat_all <- dat_all %>% 
    mutate(
        Year = lubridate::year(Date),
        Month = lubridate::month(Date),
        Day = lubridate::mday(Date),
        Date = format(Date, "%m/%d/%Y")
    ) 
```

```{r}
names(dat_all)
```

# FIX KNOWN ISSUES  

In the 2020 file, 'Lat' is repeated in the 'Long' column for Segment B, Plot 3-3. 

Per email with Tracy on 5/16/22, coordinates for this plot are:  
33.34045339	-79.2030272  

Code used to fix it in the mapping script was:  

```{r}
# file_in <- here::here(path_start, "NIW", "NIWVEG2020.xlsx")
# niw <- read_cdmo(file_in) %>% 
#     mutate(id = paste(SiteID, TransectID, PlotID, sep = "-"),
#            Lat = case_when(id == "Segment B-3-3" ~ 33.34045339,
#                            TRUE ~ Lat),
#            Long = case_when(id == "Segment B-3-3" ~ -79.2030272,
#                             TRUE ~ Long))
```

Fix here:  

```{r}
dat_all <- dat_all %>% 
    mutate(id = paste(SiteID, TransectID, PlotID, sep = "-"),
           Lat = case_when(id == "Segment B-3-3" ~ 33.34045339,
                           TRUE ~ Lat),
           Long = case_when(id == "Segment B-3-3" ~ -79.2030272,
                            TRUE ~ Long)) %>% 
    select(-id)
```


## Species names fixes  

```{r}
dat_all <- dat_all %>% 
    mutate(Species = case_when(Species == "Salicornia biglovii" ~ "Salicornia bigelovii",
                               Species == "Juniperus virginicus" ~ "Juniperus virginiana",
                               Species == "Fimbristylus spadicea" ~ "Fimbristylis spadicea",
                               TRUE ~ Species))
```

## Subplot fix  

For Segment A, Transect 3, Plot 4, unvegetated, there's a 'subplot' ID of 3 but only cover (not density or height) is provided - more consistent with 'NA' for subplot. Changing here.  

```{r}
dat_all <- dat_all %>% 
    mutate(long_id = paste(SiteID, TransectID, PlotID, Species, sep = "-"),
           Subplot = case_when(long_id == "Segment A-3-4-unvegetated" ~ NA_real_,
                               TRUE ~ Subplot)) %>% 
    select(-long_id)
```


## CDMO format (long)  

CDMO column names, in order. If columns had different names, or there were additional ones, account for that here.    

```{r}
dat_cdmo <- dat_all %>% 
    select(
        "Reserve",
        "Type",
        "Date",
        "Year",
        "Month",
        "Day",
        "SiteID",
        "TransectID",
        "PlotID",
        "Subplot",
        "Rep",
        "SSAM-1",
        "Lat",
        "Long",
        "Distance",
        "Orthometric_Height",
        "Height_Relative_to_MLLW",
        "Species",
        "Cover",
        "Density",
        "Maximum Canopy Height",
        "Average Canopy Height",
        "Diameter",
        "Height",
        "QAQC"
    )
```

## Quality Checks  

```{r}
dat_cdmo %>% 
    select(Reserve:Species, QAQC) %>%
    janitor::get_dupes() %>% 
    View()
```

```{r}
dat_cdmo %>% 
    select(Reserve, SiteID, TransectID, PlotID,
           Distance) %>% 
    distinct() %>% 
    janitor::get_dupes(-Distance)
```



## Correct discrepancies  

A few plots had lat/long/elevation errors (copied from the plot above them) that made extra rows in the NMST format. Those time periods were:  

+  09/14/2010, Segment B, Transect 2, Plot 9  
+  09/18/2007, Segment B, Transect 1, Plot 5  
+  09/22/2015, Segment B, Transect 2, Plot 9  

Want to replace these with the values from the row below each (from subplot 1 into subplot 'NA').  

Date is character format, so it ought to be easy to pull out the indices and do it.  

Update 11/28/2023: Additional problems, that caused duplicate rows in the Density sheet:  

+  08/27/2009, Segment B, Transect 1, Plot 2  
+  08/28/2009, Segment A, Transect 2, Plot 1  

```{r}
# pull out column indices
colsToFix <- which(names(dat_cdmo) %in% c("SSAM-1", "Lat", "Long", "Distance", "Orthometric_Height", "Height_Relative_to_MLLW"))


# FIRST IN THE LIST
# pull out row indices
issue1 <- dat_cdmo %>% 
    mutate(rownum = row_number()) %>% 
    filter(Date == "09/14/2010",
           SiteID == "Segment B",
           TransectID == 2,
           PlotID == 9) %>% 
    select(rownum) %>% 
    unlist()
# make sure I'm looking at the right things
dat_cdmo[issue1, colsToFix]
# make the change
dat_cdmo[issue1[1], colsToFix] <- dat_cdmo[issue1[2], colsToFix]
# make sure it looks fixed
dat_cdmo[issue1, colsToFix]

# SECOND IN THE LIST
# pull out row indices
issue2 <- dat_cdmo %>% 
    mutate(rownum = row_number()) %>% 
    filter(Date == "09/18/2007",
           SiteID == "Segment B",
           TransectID == 1,
           PlotID == 5) %>% 
    select(rownum) %>% 
    unlist()

# make sure I'm looking at the right things
dat_cdmo[issue2, colsToFix]
# make the change
dat_cdmo[issue2[1], colsToFix] <- dat_cdmo[issue2[2], colsToFix]
# make sure it looks fixed
dat_cdmo[issue2, colsToFix]


# THIRD IN THE LIST
# pull out row indices
issue3 <- dat_cdmo %>% 
    mutate(rownum = row_number()) %>% 
    filter(Date == "09/22/2015",
           SiteID == "Segment B",
           TransectID == 2,
           PlotID == 9) %>% 
    select(rownum) %>% 
    unlist()

# make sure I'm looking at the right things
dat_cdmo[issue3, colsToFix]
# make the change
dat_cdmo[issue3[1], colsToFix] <- dat_cdmo[issue3[2], colsToFix]
# make sure it looks fixed
dat_cdmo[issue3, colsToFix]



# 4th IN THE LIST
# pull out row indices
issue4 <- dat_cdmo %>% 
    mutate(rownum = row_number()) %>% 
    filter(Date == "08/27/2009",
           SiteID == "Segment B",
           TransectID == 1,
           PlotID == 2) %>% 
    select(rownum) %>% 
    unlist()

# the first and last of those rows are wonky; replace with the 2nd
# see other rows from this site-transect-plot combination to verify
# issue4b <- dat_cdmo %>% 
#     mutate(rownum = row_number()) %>% 
#     filter(SiteID == "Segment B",
#            TransectID == 1,
#            PlotID == 2) 

# make sure I'm looking at the right things
dat_cdmo[issue4, colsToFix]
# make the changes
dat_cdmo[issue4[1], colsToFix] <- dat_cdmo[issue4[2], colsToFix]
dat_cdmo[issue4[4], colsToFix] <- dat_cdmo[issue4[2], colsToFix]
# make sure it looks fixed
dat_cdmo[issue4, colsToFix]


# 5th IN THE LIST
# pull out row indices
issue5 <- dat_cdmo %>% 
    mutate(rownum = row_number()) %>% 
    filter(Date == "08/28/2009",
           SiteID == "Segment A",
           TransectID == 2,
           PlotID == 1) %>% 
    select(rownum) %>% 
    unlist()

# the ones with ortho ht around 1 are the correct ones
# see other rows from this site-transect-plot combination to verify
# issue5b <- dat_cdmo %>%
#     mutate(rownum = row_number()) %>%
#     filter(SiteID == "Segment A",
#            TransectID == 2,
#            PlotID == 1)

# make sure I'm looking at the right things
dat_cdmo[issue5, colsToFix]
# make the changes
dat_cdmo[issue5[c(1, 5, 6, 10)], colsToFix] <- dat_cdmo[issue5[2], colsToFix]
# make sure it looks fixed
dat_cdmo[issue5, colsToFix]
```




## Write the CDMO data frame to a file. **Modify for each reserve**  

```{r}
write_xlsx(dat_cdmo, path = here::here(path_out, "NIW_CDMO.xlsx"),
           format_headers = TRUE)
```


# NaMaSTe tables  

Get rid of any NAs in lat/long; we want station characteristics.  

```{r}
station_table <- dat_cdmo %>% 
    select(Reserve,
           SiteID,
           TransectID,
           PlotID,
           Lat,
           Long,
           Type,
           "SSAM-1") %>% 
    filter(!is.na(Lat),
           !is.na(Long)) %>% 
    distinct()
```

Do any plots have multiple rows?  

```{r}
station_table %>% 
    group_by(Reserve, SiteID, TransectID, PlotID) %>% 
    tally() %>% 
    arrange(desc(n)) %>% 
    knitr::kable()
```

Are all stations actually represented? If the below output is FALSE, there's a problem to fix. If TRUE, proceed.     

```{r}
stns_all <- dat_cdmo %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct()

nrow(stns_all) == nrow(station_table)
```


# NIW to Namaste-ish  

Need to pull out subplots for density and height.  

The following isn't really working. From the metadata, subplots are only done for density, and only high-density species. So what I really need to do is, pull out all the density data, and pivot that separately from everything else......

```{r}
covr_plots <- dat_cdmo %>% 
    filter(is.na(Subplot))
# should have values for cover; NAs for density and height
summary(covr_plots[c("Cover", "Density", "Maximum Canopy Height", "Average Canopy Height")])

sub_plots <- dat_cdmo %>% 
    filter(!is.na(Subplot))
summary(sub_plots[c("Cover", "Density", "Maximum Canopy Height", "Average Canopy Height")])
```

```{r}
dens_sub <- dat_cdmo %>% 
    filter(!is.na(Density)) %>% 
    select(Reserve:Species, Density, QAQC) %>% 
    mutate(Subplot = case_when(is.na(Subplot) ~ 1,
                               TRUE ~ Subplot),
           Species = paste(Species, Subplot, sep = " "))


ht_sub <- dat_cdmo %>% 
    filter(!is.na(`Maximum Canopy Height`)) %>% 
    select(Reserve:Species, `Maximum Canopy Height`, QAQC) 

qaqc_to_join <- dat_cdmo %>% 
    filter(is.na(Subplot)) %>% 
    select(Reserve:Species, QAQC) %>% 
    pivot_wider(names_from = Species, 
                values_from = QAQC, 
                names_prefix = "F_",
                names_sort = TRUE)
```

Pivot to wide  


```{r}
dens_wide <- dens_sub %>% 
    select(-QAQC, -Subplot) %>% 
    pivot_wider(names_from = Species, 
                values_from = Density,
                names_sort = TRUE)

ht_wide <- ht_sub %>% 
    select(-QAQC, -Subplot) %>% 
    pivot_wider(names_from = Species, 
                values_from = `Maximum Canopy Height`,
                names_sort = TRUE)

covr_wide <- dat_cdmo %>% 
    filter(is.na(Subplot)) %>% 
    select(-Subplot, -c(Density:QAQC)) %>% 
    pivot_wider(names_from = Species, 
                values_from = Cover,
                names_sort = TRUE)
```

Need to average whatever values are present for density, then make it wide again, then join to cover and height. Need to add 'density' or 'maximum canopy height' as prefixes.  Actually don't need the acrobatics until future work, when data might be entered into the density sheet. Here, can just work with dens_sub.  

```{r}
dens_means <- dat_cdmo %>% 
    filter(!is.na(Density)) %>% 
    select(Reserve:Species, Density) %>% 
    group_by(Reserve, Type, Date, SiteID, TransectID, PlotID, Species) %>% 
    summarize(Density = mean(Density, na.rm = TRUE))
    
dens_to_join <- dens_means %>% 
    pivot_wider(names_from = Species, 
                values_from = Density, 
                names_prefix = "Density_",
                names_sort = TRUE)

ht_to_join <- ht_sub %>% 
    select(-QAQC, -Subplot) %>% 
    pivot_wider(names_from = Species, 
                values_from = `Maximum Canopy Height`,
                names_prefix = "Maximum Canopy Height_",
                names_sort = TRUE)
```


Join:  

```{r}
covr_sheet_out <- covr_wide %>% 
    mutate(Notes = NA_character_) %>% 
    full_join(dens_to_join) %>% 
    full_join(ht_to_join) %>% 
    full_join(qaqc_to_join) %>% 
    relocate(Notes, .after = Height_Relative_to_MLLW) %>% 
    select(-`NA`, -Subplot, -F_NA,
           -Date, -Type, -`SSAM-1`,
           -Lat, -Long, -Rep)

# write_xlsx(covr_sheet_out, path = here::here("wrangled_data", "NMST", "NIW_NMSTish.xlsx"))
```


# Other sheets  

Density and Height sheets can be built from `dens_wide` and `ht_wide` but I'm worried rows won't 100% match the cover sheet.  

```{r}
dens_sheet <- dens_wide %>% 
    select(-c(Type, Date, `SSAM-1`, Rep, Lat, Long)) %>% 
    mutate(Notes = NA_character_) %>% 
    relocate(Notes, .after = Height_Relative_to_MLLW)
hts_sheet <- ht_wide %>% 
    select(-c(Type, Date, `SSAM-1`, Rep, Lat, Long)) %>% 
    mutate(Notes = NA_character_) %>% 
    relocate(Notes, .after = Height_Relative_to_MLLW)
```



Back to code from `pivot_CDMO-to-NMST.Rmd`  

# Station Table  

Everything in this table is the first recorded value.  


```{r}
dat <- dat_cdmo 
# there must be a better way to do this, but it works.

stns <- dat %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct() %>% 
    arrange(SiteID, TransectID, PlotID)
lats <- establ(dat, Lat)
longs <- establ(dat, Long)
ssam1 <- establ(dat, `SSAM-1`)
dists <- establ(dat, Distance)
orth_hts <- establ(dat, Orthometric_Height) 
mllw_hts <- establ(dat, Height_Relative_to_MLLW)
types <- establ(dat, Type)

stn_tbl <- full_join(stns, lats) %>% 
    full_join(longs) %>% 
    full_join(types) %>% 
    full_join(orth_hts) %>% 
    full_join(mllw_hts) %>%
    full_join(dists) %>% 
    full_join(ssam1)

```



# Species List  

```{r}
spp_names <- dat %>% 
    select(Species) %>% 
    distinct() %>% 
    filter(!is.na(Species)) %>% 
    arrange(Species)
```


# Make sure dupes are gone  

```{r}
NMST_dupes <- covr_sheet_out %>% 
    mutate(NMSTrownum = row_number()) %>% 
    janitor::get_dupes(Year, Month, Day, SiteID, TransectID, PlotID) %>% 
    relocate(NMSTrownum) %>% 
    select(NMSTrownum:Height_Relative_to_MLLW)

# View(NMST_dupes)
```



# Write to Excel  

```{r}
write_xlsx(list(Station_Table = stn_tbl,
                Species_Names = spp_names,
                Cover = covr_sheet_out,
                Density = dens_sheet,
                Height = hts_sheet),
           path = out_path,
           format_headers = TRUE)
```
