---
title: "NIW"
author: "Kim Cressman"
date: "2022-05-25; latest update `r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_depth: 2
        code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

```{r}
library(tidyverse)
library(readxl)
library(writexl)
library(leaflet)
library(rgdal)
library(stringi)  # for stri_sub_replace()
```

# Setup  

## Helpful functions  

Because I'll be doing this repetitively on similar data files....  

```{r}
source(here::here("R", "sourced", "00_helper_functions.R"))
```

## Data path  

Update for each reserve: all start at  `here::here("submitted_data", "data")` but it varies by reserve.   

Also create an output folder for each reserve in the `wrangled data` folder and update the `path_out` below.  

```{r}
path_start <- here::here("submitted_data", "data", "NIW")
path_out <- here::here("wrangled_data", "NIW")
```


What are the files?  

```{r}
dir(path_start)  
```

Which ones to read in?  **MODIFY FOR EACH RESERVE**  

```{r}
step1 <- str_starts(dir(path_start), "NIWVEG")
step2 <- str_detect(dir(path_start), ".xlsx")
step3 <- step1 + step2


to_read <- dir(path_start)[which(step3 == 2)]
```

Read into a list and combine  **MODIFY FOR DIFFERENT DATA TYPES**  

# CDMO formatted input  

This section will compile code to read in and combine files that are already in CDMO format. Then a long CDMO-format file will be generated, after reconciling column names, dates, and putting things in order. The NaMaSTe template file will be generated from the CDMO-format file.  

```{r}
dat_in <- list()

for(i in seq_along(to_read)){
    dat_in[[i]] <- read_cdmo(here::here(path_start, to_read[i])) 
}

dat_all <- bind_rows(dat_in)
```


# Checks  

## Column names and types?  

Look for different column names for the same thing.  

```{r}
names(dat_all)
```

## Duplicates in date-site-transect-plot-species  

Include subplot if used by reserve. Want an empty tibble below.    

```{r}
# subplot included
dat_all %>% 
   group_by(Date, SiteID, TransectID, PlotID, Subplot, Species) %>% 
    tally() %>% 
    filter(n > 1) %>% 
    select(Date:Species, n)

# # no subplot included
# dat_all %>% 
#    group_by(Date, SiteID, TransectID, PlotID,  Species) %>% 
#     tally() %>% 
#     filter(n > 1) %>% 
#     select(Date:Species, n)
```


## Check station/plot names  

```{r}
unique(dat_all$Reserve)
dat_all %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct() %>% 
    knitr::kable()
```


Fix any issues here.  

```{r}

```


## Mangroves or SAV present?  

Mangrove species are: *Avicennia germinans*, *Laguncularia racemosa*, *Rhizophora mangle*. Check for these before discarding a `Height` or `Diameter` column.   

The following sums will be 0 if none of the genus names were found in the species column. This is only set to detect the first few letteres in case of misspellings, so if any sums are above 0, it deserves further investigation.  

```{r}
spp_all <- sort(unique(dat_all$Species))
knitr::kable(spp_all)
sum(str_detect(spp_all, "Avic"))
sum(str_detect(spp_all, "Lagunc"))
sum(str_detect(spp_all, "Rhizo"))

```

Also check for anything other than 'E' in 'Type'.

```{r}
unique(dat_all$Type)
```




## Column wrangling  

Check for columns that may have been named differently in different files.  

```{r}
names(dat_all)
```

If there are any necessary changes, make them here. Also pull out columns for Month, Day, and Year; and format Date as mm/dd/yyyy for the CDMO output file.  

```{r}
dat_all <- dat_all %>% 
    mutate(
        Year = lubridate::year(Date),
        Month = lubridate::month(Date),
        Day = lubridate::mday(Date),
        Date = format(Date, "%m/%d/%Y")
    ) 
```

```{r}
names(dat_all)
```

# FIX KNOWN ISSUES  

In the 2020 file, 'Lat' is repeated in the 'Long' column for Segment B, Plot 3-3. 

Per email with Tracy on 5/16/22, coordinates for this plot are:  
33.34045339	-79.2030272  

Code used to fix it in the mapping script was:  

```{r}
# file_in <- here::here(path_start, "NIW", "NIWVEG2020.xlsx")
# niw <- read_cdmo(file_in) %>% 
#     mutate(id = paste(SiteID, TransectID, PlotID, sep = "-"),
#            Lat = case_when(id == "Segment B-3-3" ~ 33.34045339,
#                            TRUE ~ Lat),
#            Long = case_when(id == "Segment B-3-3" ~ -79.2030272,
#                             TRUE ~ Long))
```






## CDMO format (long)  

CDMO column names, in order. If columns had different names, or there were additional ones, account for that here.    

```{r}
dat_cdmo <- dat_all %>% 
    select(
        "Reserve",
        "Type",
        "Date",
        "SiteID",
        "TransectID",
        "PlotID",
        "Subplot",
        "Rep",
        "SSAM-1",
        "Lat",
        "Long",
        "Distance",
        "Orthometric Height",
        "Height Relative to MLLW",
        "Species",
        "Cover",
        "Density",
        "Maximum Canopy Height",
        "Average Canopy Height",
        "Diameter",
        "Height",
        "QAQC"
    )
```

## Quality Checks  

```{r}
dat_cdmo %>% 
    select(Reserve:Species, QAQC) %>%
    janitor::get_dupes() %>% 
    View()
```

```{r}
dat_cdmo %>% 
    select(Reserve, SiteID, TransectID, PlotID,
           Distance) %>% 
    distinct() %>% 
    janitor::get_dupes(-Distance)
```


## Correct discrepancies  


## Write the CDMO data frame to a file. **Modify for each reserve**  

```{r}
write_xlsx(dat_cdmo, path = here::here(path_out, "Reserve_CDMO.xlsx"),
           format_headers = TRUE)
```


# NaMaSTe tables  

Get rid of any NAs in lat/long; we want station characteristics.  

```{r}
station_table <- dat_cdmo %>% 
    select(Reserve,
           SiteID,
           TransectID,
           PlotID,
           Lat,
           Long,
           Type,
           "SSAM-1") %>% 
    filter(!is.na(Lat),
           !is.na(Long)) %>% 
    distinct()
```

Do any plots have multiple rows?  

```{r}
station_table %>% 
    group_by(Reserve, SiteID, TransectID, PlotID) %>% 
    tally() %>% 
    arrange(desc(n)) %>% 
    knitr::kable()
```

Are all stations actually represented? If the below output is FALSE, there's a problem to fix. If TRUE, proceed.     

```{r}
stns_all <- dat_cdmo %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct()

nrow(stns_all) == nrow(station_table)
```