---
title: "SOS to NMST"
author: "Kim Cressman"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

# Import data  

```{r}
library(tidyverse)
library(readxl)
library(writexl)
library(leaflet)
library(rgdal)
library(stringi)  # for stri_sub_replace()


# function to pick first non-NA reading for any given plot ("at establishment")
establ <- function(df, x){
    
    x_vec <- df %>% 
        select({{x}})
    
    if(sum(!is.na(x_vec)) == 0){
        out <- df %>% 
            select(SiteID, TransectID, PlotID, {{x}}) %>% 
            distinct() 
    } else {
        out <- df %>% 
            select(Date, SiteID, TransectID, PlotID, {{x}}) %>% 
            distinct() %>% 
            group_by(SiteID, TransectID, PlotID) %>% 
            filter(!is.na({{x}})) %>%
            filter(Date == min(Date)) %>% 
            select(-Date)
    }
        
    out
}



reserve <- "SOS"
```

```{r}
# path <- here::here("wrangled_data", "CDMO", "res_CDMO.xlsx")
path <- here::here("wrangled_data", "CDMO", paste0(reserve, "_CDMO.xlsx"))

# out_path <- here::here("wrangled_data", "NMST", "res_NMST.xlsx")
out_path <- here::here("wrangled_data", "NMST", paste0(reserve, "_NMST.xlsx"))
```

Renaming possibilities:  

```{r}
col_matching <- c("Orthometric_Height" = "Orthometric Height",
                  "Height_Relative_to_MLLW" = "Height Relative to MLLW",
                  "Ht" = "Canopy Height",
                  "Ht" = "Average Canopy Height",
                  "Ht" = "Maximum Canopy Height")

# deal with different types of canopy height measurement
res_info <- read_xlsx(here::here("Reserve_Info.xlsx"))
ht_type <- unlist(res_info[res_info$Reserve == reserve, "Height_CDMO"])
```

```{r}
dat <- read_xlsx(path,
                 guess_max = 5000,
                 na = c("", "NA", "N/A", "na")) %>% 
    rename(any_of(col_matching)) %>% 
    filter(!is.na(Year)) %>% 
    mutate(Date = lubridate::ymd(paste(Year, Month, Day, sep = "-")))
```

Add required columns if they don't exist. QAQC, SSAM-1, OrthoHt, HtMLLW.

```{r}
if(!exists("QAQC", dat)){
  dat$QAQC <- NA  
}

if(!exists("SSAM-1", dat)){
  dat$`SSAM-1` <- NA  
}

if(!exists("Orthometric_Height", dat)){
  dat$Orthometric_Height <- NA  
}

if(!exists("Height_Relative_to_MLLW", dat)){
  dat$Height_Relative_to_MLLW <- NA  
}
```


Make sure every row has the appropriate value for Date_HtDens. Sort by site-transect-plot and Date_HtDens, then fill.   

```{r}
dat <- dat %>% 
    arrange(Year, SiteID, TransectID, PlotID, Date_HtDens) %>% 
    fill(Date_HtDens, .direction = "down")

# make sure I didn't just introduce duplicates or weirdnesses  
# dat2 %>% 
#     select(Year, SiteID, TransectID, PlotID, Date_HtDens) %>% 
#     distinct() %>% 
#     janitor::get_dupes()
```

Because of some issues writing out the NMST file, I'm formatting both extra date columns as character.  

```{r}
dat <- dat %>% 
    mutate(Date_HtDens = as.character(Date_HtDens),
           Date_GPS = as.character(Date_GPS))
```



## Identifiers  

Should be the same between all sheets  

update on 1/12/2023: added "Date_HtDens" to id_names  
3/8/23: aded "Date_GPS" to id_names  

```{r}
id_names <- c("Reserve", "SiteID", 
             "Year", "Month", "Day",
             "Date_HtDens", "Date_GPS",
             "TransectID", "PlotID", # "Subplot",
             "Orthometric_Height", "Height_Relative_to_MLLW",
             "Distance")

# only use ID columns that exist in the data frame
id_names <- id_names[which(id_names %in% names(dat))]
```

## Deal with notes separately  

If there's not a Notes column, make one full of NAs.  

NOTE there was a notes column in the original data files but it was not propagated here.....

```{r}
if(exists("Notes", dat)){
    notes <- dat %>% 
        filter(!is.na(Notes))
    c_notes <- notes %>% 
        group_by(SiteID, Year, Month, Day, TransectID, PlotID) %>% 
        summarize(Notes = str_c(unique(Notes), collapse = "; "))
} else {
    c_notes <- dat %>% 
        group_by(SiteID, Year, Month, Day, TransectID, PlotID) %>% 
        summarize(Notes = NA_character_)
}
```

# Cover sheet  

NOTE this sheet also has average canopy height and density. Replicates are on the other sheets.  

**For SOS, still using code from `pivot_CDMO-to-NMST.Rmd` for this sheet** - will re-import full density and height data for the density and height sheets.  

## Cover data  

```{r}
covr <- dat %>% 
    select(all_of(id_names), Species, Cover, QAQC)

# make sure there aren't any dupes  
janitor::get_dupes(covr, -Cover)
```

```{r}
# pivot
covr_wide <- covr %>% 
    rename("F" = QAQC) %>% 
    pivot_wider(names_from = Species,
                values_from = c(Cover, "F"),
                names_sort = TRUE)

# get the 'Cover' prefix out of the names
names(covr_wide) <- str_replace(names(covr_wide), "^(Cover_)", "")

# get rid of the NA and F_NA columns
covr_wide <- covr_wide %>% 
    select(-"NA", -"F_NA")

# check again for dupes
janitor::get_dupes(covr_wide, Year, Month, Day, SiteID, TransectID, PlotID)
```


## Density data  

```{r}
dens <- dat %>% 
    select(all_of(id_names), Species, Density) %>% 
    filter(!is.na(Density)) %>% 
    pivot_wider(names_from = Species,
                values_from = Density,
                names_prefix = "Density_",
                names_sort = TRUE)
```


## Height data  

```{r}
hts <- dat %>% 
    select(all_of(id_names), Species, Ht) %>% 
    filter(!is.na(Ht)) %>% 
    pivot_wider(names_from = Species,
                values_from = Ht,
                names_prefix = paste0(ht_type, "_"),
                names_sort = TRUE)
```


## Mangrove height and diameter  

if applicable.  

```{r}
spp_all <- sort(unique(dat$Species))

# are mangroves present in the species list
mangroves <- sum(str_detect(spp_all, "Avic")) + sum(str_detect(spp_all, "Lagunc")) + sum(str_detect(spp_all, "Rhizo"))

# what are those species
mang_spp <- spp_all[str_which(spp_all, "Avic")]

if(mangroves > 0){
    dat_mang_ht <- dat %>% 
        filter(Species %in% mang_spp) %>% 
        select(all_of(id_names), Species, Height) %>% 
        pivot_wider(names_from = Species,
                values_from = Height,  # Height in original dataset is mangrove height
                names_prefix = "Height_")
    
    dat_mang_diam <- dat %>% 
        filter(Species %in% mang_spp) %>% 
        select(all_of(id_names), Species, Diameter) %>% 
        pivot_wider(names_from = Species,
                values_from = Diameter,  # Height in original dataset is mangrove height
                names_prefix = "Diameter_")
    
    dat_mang <- full_join(dat_mang_ht, dat_mang_diam)
}
```


## Join them all  

```{r}
covr_sheet <- full_join(covr_wide, dens) %>% 
    full_join(hts) %>% 
    full_join(c_notes) %>% 
    select(all_of(id_names), Notes, everything()) %>% 
    relocate(Notes, .after = Distance) %>% 
    relocate(starts_with("F_"), .after = last_col())

# join in mangrove heights and diameters too
if(mangroves > 0){
    covr_sheet <- full_join(covr_sheet, dat_mang) %>% 
        relocate(starts_with("F_"), .after = last_col())
}
```


# Density and Height Sheets  

## Bring in density and heights workbook  

```{r}
path_start <- here::here("submitted_data", "data", reserve)
dens_ht <- read_xlsx(here::here(path_start, "SOS_StemCt_Ht_2010_2021.xlsx"),
                     sheet = "Stem_counts_and_heights")
abbrevs <- read_xlsx(here::here(path_start, "SOS_StemCt_Ht_2010_2021.xlsx"),
                     sheet = "Plant Acronyms")
names(abbrevs) <- c("code", "Species")
```

Issues to fix:  

+  Date is sometimes mm/dd/yyyy and other times Excel's date format. Split these out and unify.  
+  Site DP is really DM; corrected here  

```{r}
dens_ht <- dens_ht %>% 
    select(-Year) %>% 
    mutate(Date_excel = as.numeric(case_when(!str_detect(Date, "/") ~ Date,
                                             TRUE ~ NA_character_)),
           Date_mdy = case_when(str_detect(Date, "/") ~ Date,
                                TRUE ~ NA_character_),
           Date = case_when(!is.na(Date_excel) ~ janitor::excel_numeric_to_date(Date_excel),
                            !is.na(Date_mdy) ~ lubridate::mdy(Date_mdy)),
           across(c(`AGRSTO ct`:`OENSAR ht3`), as.numeric),
           TransectID = as.character(TransectID),
           PlotID = as.character(PlotID),
           Site = case_when(Site == "DP" ~ "DM",
                            TRUE ~ Site),
           Year = lubridate::year(Date)) %>% 
    rename(SiteID = Site) %>% 
    relocate(Year, .after = Date) %>% 
    select(-Date_excel, -Date_mdy)
```

Added Year column for joining with cover sheet. Some dates possibly still wrong in ht/dens spreadsheet so really only want to join on year-site-transect-plot and pull the additional info from the cover data frame.  

## Split out density and height  

Pivoted longer to replace codes with names, then pivot back wider. Probably don't need to do it this way but the pivot_longer code is already written from generating the averages for the cover sheet.  

```{r}
all_names <- names(dens_ht)
dens_names <- all_names[str_ends(all_names, pattern = " ct")]
ht_names <- all_names[str_ends(all_names, pattern = " ht[1-3]")]

dens <- dens_ht %>% 
    select(ID:Year, "Notes" = NOTES, all_of(dens_names)) %>% 
    pivot_longer(-(ID:Notes), 
                 names_to = c("code", "param"), 
                 names_sep = " ",
                 values_to = "Density") %>% 
    filter(!is.na(Density)) %>% 
    full_join(abbrevs) %>% 
    select(-code, -param) %>% 
    relocate(Species, .before = Density)
    
hts <- dens_ht %>% 
    select(ID:Year, "Notes" = NOTES, all_of(ht_names))%>% 
    pivot_longer(-(ID:Notes), 
                 names_to = c("code", "rep"), 
                 names_sep = " ht",
                 values_to = "Height") %>% 
    filter(!is.na(Height)) %>%
    full_join(abbrevs) %>% 
    select(-code) %>% 
    relocate(Species, .before = Height) %>%
    mutate(Species = paste(Species, rep)) %>% 
    select(-rep)
```

## Pivot  

```{r}
dens_wide <- dens %>% 
    select(SiteID, TransectID, PlotID, Year, Species, Density) %>% 
    filter(!is.na(Density)) %>% 
    pivot_wider(names_from = Species,
                values_from = Density,
                names_sort = TRUE)

hts_wide <- hts %>% 
    select(SiteID, TransectID, PlotID, Year, Species, Height) %>%
    filter(!is.na(Height)) %>% 
    pivot_wider(names_from = Species,
                values_from = Height,
                names_sort = TRUE)
```

Join back with ID columns from cover data frame  

```{r}
dens_sheet <- covr_sheet %>% 
    select(Reserve:Notes) %>% 
    full_join(dens_wide, by = c("Year", "SiteID", "TransectID", "PlotID"))

hts_sheet <- covr_sheet %>% 
    select(Reserve:Notes) %>% 
    full_join(hts_wide, by = c("Year", "SiteID", "TransectID", "PlotID"))
```


# Other sheets  

Back to code from `pivot_CDMO-to-NMST.Rmd`  

# Station Table  

Everything in this table is the first recorded value.  


```{r}
# there must be a better way to do this, but it works.

stns <- dat %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct() %>% 
    arrange(SiteID, TransectID, PlotID)
lats <- establ(dat, Lat)
longs <- establ(dat, Long)
ssam1 <- establ(dat, `SSAM-1`)
dists <- establ(dat, Distance)
orth_hts <- establ(dat, Orthometric_Height) 
mllw_hts <- establ(dat, Height_Relative_to_MLLW)
types <- establ(dat, Type)

stn_tbl <- full_join(stns, lats) %>% 
    full_join(longs) %>% 
    full_join(types) %>% 
    full_join(orth_hts) %>% 
    full_join(mllw_hts) %>%
    full_join(dists) %>% 
    full_join(ssam1) %>% 
    arrange(SiteID, TransectID, PlotID)

```

Some DM sites don't have associated lat/long/elevation data. There were several sites with PlotID listed as NA; wonder if these belong to the missing plot numbers?  


# Species List  

```{r}
spp_names <- dat %>% 
    select(Species) %>% 
    distinct() %>% 
    arrange(Species)
```




# Write to Excel  

```{r}
write_xlsx(list(Station_Table = stn_tbl,
                Species_Names = spp_names,
                Cover = covr_sheet,
                Density = dens_sheet,
                Height = hts_sheet),
           path = out_path,
           format_headers = TRUE)
```

