---
title: "TJR"
author: "Kim Cressman"
date: "2023-01-30; latest update `r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_depth: 2
        code_folding: hide
---

Use this template to set up the wrangling for an individual reserve's files.  


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

```{r}
library(tidyverse)
library(readxl)
library(writexl)
library(leaflet)
library(rgdal)
library(stringi)  # for stri_sub_replace()

reserve <- "TJR"
```

# Setup  

## Helpful functions  

Because I'll be doing this repetitively on similar data files....  

```{r}
source(here::here("R", "sourced", "00_helper_functions.R"))
```

## Data path  

Update for each reserve: all start at  `here::here("submitted_data", "data")` but it varies by reserve.   

Also create an output folder for each reserve in the `wrangled data` folder and update the `path_out` below.  

```{r}
path_start <- here::here("submitted_data", "data", reserve)
problem_path_out <- here::here("wrangled_data", "combined_with_issues")
path_out <- here::here("wrangled_data", "CDMO")

# create the output folder if it doesn't already exist
if(!dir.exists(path_out)){
    dir.create(path_out)
}
```


What are the files?  

```{r}
dir(path_start)  
```

Which ones to read in?  **MODIFY FOR EACH RESERVE**  

```{r}
file_in <- here::here(path_start, "All TJE VEG 15Nov22 v2 for NAMASTE.xlsx")

dat_all <- read_xlsx(file_in) %>% 
    rename(SiteID = Site,
           TransectID = Transect,
           PlotID = Quadrat,
           Cover = PercentCover)
```


# Checks  

## Column names and types?  

Look for different column names for the same thing.  

```{r}
names(dat_all)
```

## Duplicates in date-site-transect-plot-species  

Include subplot if used by reserve. Want an empty tibble below.    

```{r}
# unique(dat_all$Subplot)

# # subplot included
# dat_all %>% 
#    group_by(Date, SiteID, TransectID, PlotID, Subplot, Species) %>% 
#     tally() %>% 
#     filter(n > 1) %>% 
#     select(Date:Species, n)

# no subplot included
dat_all %>%
   group_by(Year, SiteID, TransectID, PlotID,  Species) %>%
    tally() %>%
    filter(n > 1) %>%
    select(Year:Species, n)

TJR_dupes <- janitor::get_dupes(dat_all,
                   Year, Season, SiteID, TransectID, PlotID, Species,
                   "Live/Dead") 

write_xlsx(TJR_dupes, here::here(problem_path_out, 
                                 "TJR_dupes.xlsx"))
```


## Check station/plot names  

```{r}
dat_all$Reserve <- "TJR"
unique(dat_all$Reserve)
```

```{r}
dat_all %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct() %>% 
    arrange(SiteID, TransectID, PlotID) %>% 
    knitr::kable()
```


### 'Distance' discrepancies  

no Distance column

```{r}
# stn_dupes <- dat_all %>% 
#     select(Reserve, SiteID, TransectID, PlotID,
#            Distance) %>% 
#     distinct() %>% 
#     janitor::get_dupes(-Distance)
# 
# stn_dupes
```


### Lat/Long discrepancies  

no lat/long in data


## Mangroves or SAV present?  

Mangrove species are: *Avicennia germinans*, *Laguncularia racemosa*, *Rhizophora mangle*. Check for these before discarding a `Height` or `Diameter` column.   

The following sums will be 0 if none of the genus names were found in the species column. This is only set to detect the first few letteres in case of misspellings, so if any sums are above 0, it deserves further investigation.  

```{r}
spp_all <- sort(unique(dat_all$Species))
knitr::kable(spp_all)
sum(str_detect(spp_all, "Avic"))
sum(str_detect(spp_all, "Lagunc"))
sum(str_detect(spp_all, "Rhizo"))

```

Species names all look okay!

Also check for anything other than 'E' in 'Type'.

```{r}
dat_all$Type <- "E"
unique(dat_all$Type)
```

## Check species names  

```{r}
dat_all$Species <- str_replace(dat_all$Species, pattern = "  ", replacement = " ")
    
spp <- dat_all %>% 
    group_by(Species) %>% 
    tally()

spp_out_path <- here::here("wrangled_data", "combined_with_issues", paste0(reserve, "_species.csv")) 

write.csv(spp, spp_out_path, row.names = FALSE)
```

# Fix discrepancies  

## Dupes  

Annotated spreadsheet says what to do.  

```{r}
dupe_fix <- read_xlsx(here::here("wrangled_data",
                                 "combined_with_issues",
                                 "TJR_dupes - with comments.xlsx"),
                      n_max = 13)

dat_all <- dat_all %>%
    full_join(dupe_fix) %>% 
    mutate(Species = case_when(!is.na(Fix_column) ~ Fix_column,
                               TRUE ~ Species)) %>% 
    filter(Species != "remove") %>% 
    select(-dupe_count, -Fix_column)


# double check for more dupes  
janitor::get_dupes(dat_all, SiteID, TransectID, PlotID, Species, `Live/Dead`, `Season/Year`)
```



## Column wrangling  

### Dates  

Pull out columns for Month, Day, and Year; and format Date as mm/dd/yyyy for the CDMO output file.  

TJR had issues with date entry so cannot provide full information at this time. For Namaste project, because date itself doesn't really matter, we will use October 15th as Month-Day for all readings. Reserve will correct after-the-fact as they are able to find time.  

```{r}
dat_all <- dat_all %>% 
    mutate(
        Month = 10,
        Day = 15,
        Date = lubridate::ymd(paste(Year, Month, Day, sep = "-")),
        Date = format(Date, "%m/%d/%Y")
    ) 
```

### Live/Dead  





## CDMO format (long)  

CDMO column names, in order. If columns had different names, or there were additional ones, account for that here.    

```{r}
dat_cdmo <- dat_all %>% 
    select(
        "Reserve",
        "Type",
        "Date",
        "Year",
        "Month",
        "Day",
        "SiteID",
        "TransectID",
        "PlotID",
        "Species",
        "Cover"
    )
```


## Write the CDMO data frame to a file. **Modify for each reserve**  

```{r}
write.csv(dat_cdmo, here::here(problem_path_out, paste0(reserve, "_issues.csv")), 
          row.names= FALSE,
          na = "")
```

```{r}
write_xlsx(dat_cdmo, path = here::here(path_out, paste0(reserve, "_CDMO.xlsx")),
           format_headers = TRUE)
```


# NaMaSTe tables  

Get rid of any NAs in lat/long; we want station characteristics.  

```{r}
station_table <- dat_cdmo %>% 
    select(Reserve,
           SiteID,
           TransectID,
           PlotID,
           Lat,
           Long,
           Type,
           "SSAM-1") %>% 
    filter(!is.na(Lat),
           !is.na(Long)) %>% 
    distinct()
```

Do any plots have multiple rows?  

```{r}
station_table %>% 
    group_by(Reserve, SiteID, TransectID, PlotID) %>% 
    tally() %>% 
    arrange(desc(n)) %>% 
    knitr::kable()
```

Are all stations actually represented? If the below output is FALSE, there's a problem to fix. If TRUE, proceed.     

```{r}
stns_all <- dat_cdmo %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct()

nrow(stns_all) == nrow(station_table)
```