---
title: "Pivoting and re-joining - GTM and NIW"
author: "Kim Cressman"
date: "5/6/2022"
output: 
    html_document:
        toc: true
        code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

```{r}
library(tidyverse)
library(readxl)

# helper functions
source(here::here("R", "sourced", "00_helper_functions.R"))

# starting path for data
path_start <- here::here("submitted_data", "data")

grouping_cols <- c("Reserve", "SiteID", "TransectID", "PlotID", "Subplot", "Date")
grouping_cols2 <- c("Reserve", "SiteID", "TransectID", "PlotID", "Date")
other_id_cols <- c("Lat", "Long", "Distance", "Orthometric Height",
                   "Height Relative to MLLW")
```

# NIW  

In this file, 'Lat' is repeated in the 'Long' column for Segment B, Plot 3-3. Removing these before binding coords because it's only one plot; will need to fix before doing other analyses.  

```{r}
file_in <- here::here(path_start, "NIW", "NIWVEG2020.xlsx")
niw <- read_cdmo(file_in) %>% 
    filter(Long < 0)
```


We have files from 2006-2020, all starting with `NIWVEG` and ending with `.xlsx`.  

```{r}
niw_list <- list()

# identify files to read - all of this condensed into the command below
## all_files <- dir(here::here(path_start, "NIW"))
## veg_files <- str_starts(all_files, "NIWVEG")
## xlsx_files <- str_ends(all_files, ".xlsx")
## veg_xlsx <- which(veg_files + xlsx_files == 2)
## file_names <- all_files[veg_xlsx]

res <- "NIW"

all_files <- dir(here::here(path_start, res))
file_names <- all_files[which(str_starts(all_files, paste0(res, "VEG")) + str_ends(all_files, ".xlsx") == 2)]

for(i in seq_along(file_names)){
    to_read <- file_names[i]
    dat <- read_cdmo2(here::here(path_start, "NIW", to_read))
    niw_list[[i]] <- dat
}

niw_long <- bind_rows(niw_list)
unique(niw_all$Species)

# get rid of species that are NA - some plots weren't sampled in 2015 because of boardwalk damage, so rows are there and there's a QAQC flag/code of -2 CSM (giving that explanation)  

niw_long <- filter(niw_long, !is.na(Species))

# # how about QAQC codes? There are some -3 CSM and 1 CSMs in there
# # are these used by species?
# niw_long %>% 
#     group_by(SiteID, TransectID, PlotID, Date) %>% 
#     summarize(n_species = length(unique(Species)),
#               n_QAQC_codes = length(unique(QAQC))) %>% 
#     arrange(desc(n_QAQC_codes),
#             desc(n_species))

# yes, they are, hm  
```

## QAQC FOR THESE PURPOSES  

Because these will be example analyses, I'm going to exclude suspect and rejected data points without looking further - this is not what I want to do overall! Just in a rush right now.  

```{r}
excl <- c("<1> (CSM)", "<-3> (CSM)", "<1>  (CSM)")
niw_qaqc <- niw_long %>% 
    filter(!(QAQC %in% excl))

# niw_wide <- niw_qaqc %>% 
#     select(-QAQC, -Height, -Diameter) %>% 
#     pivot_wider(id_cols = all_of(grouping_cols),
#                 names_from = Species,
#                 values_from = Cover:'Average Canopy Height')
```

Split up data frames and pivot. For NIW, am averaging everything from a single plot together, disregarding NAs. This should average the three replicates for height, and single value for cover, etc.    

```{r}
niw_cover <- niw_qaqc %>% 
    select(all_of(grouping_cols), Species, Cover) %>% 
    pivot_wider(names_from = Species,
                values_from = Cover)

niw_density <- niw_qaqc %>% 
    select(all_of(grouping_cols), Species, Density) %>% 
    filter(!is.na(Density)) %>% 
    mutate(Species = paste0("Density_", Species)) %>% 
    pivot_wider(names_from = Species,
                values_from = Density)

niw_height <- niw_qaqc %>% 
    select(all_of(grouping_cols), Species, 'Maximum Canopy Height') %>% 
    filter(!is.na('Maximum Canopy Height')) %>%
    mutate(Species = paste0("Maximum Canopy Height_", Species)) %>% 
    pivot_wider(names_from = Species,
                values_from = 'Maximum Canopy Height')

niw_other_id <- niw_qaqc %>% 
    select(all_of(grouping_cols), all_of(other_id_cols)) %>% 
    distinct()

niw_all <- full_join(niw_other_id, niw_cover,
                     by = all_of(grouping_cols)) %>% 
    full_join(., niw_density,
              by = all_of(grouping_cols)) %>% 
    full_join(., niw_height,
              by = all_of(grouping_cols)) %>% 
    select(-Subplot) %>% 
    group_by(Date, Reserve, SiteID, TransectID, PlotID) %>% 
    summarize(across(Lat:"Maximum Canopy Height_Ipomoea sagittata",
              ~mean(.x, na.rm = TRUE))) %>% 
    ungroup() %>% 
    select(Reserve, SiteID, TransectID, PlotID, Date, 
           all_of(other_id_cols), 
           everything())
        
    

write.csv(niw_all, here::here("data", "niw_namaste_template.csv"),
          row.names = FALSE,
          na = "")
```



# GTM  

2015 data file has an extra "Sheet1" at the beginning. Would just delete but I don't want to have to modify original files downloaded from the google drive.  

For GTM's files, the worksheet name matches the file name, so I'll use that.  

```{r}
# file_in <- here::here(path_start, "GTM", "GTMVEG2020.xlsx")
# gtm <- read_cdmo(file_in)
```

```{r}
gtm_list <- list()

res <- "GTM"

all_files <- dir(here::here(path_start, res))
file_names <- all_files[which(str_starts(all_files, paste0(res, "VEG")) + str_ends(all_files, ".xlsx") == 2)]

for(i in seq_along(file_names)){
    to_read <- file_names[i]
    sheets <- excel_sheets(here::here(path_start, res, to_read))
    sheet_to_read <- str_subset(sheets, "GTM")
    dat <- read_cdmo2(here::here(path_start, res, to_read),
                      worksheet = sheet_to_read) %>% 
        mutate(PlotID = as.numeric(PlotID))  # stored as character until 2020
    gtm_list[[i]] <- dat
    print(paste("finished", file_names[i]))
}



gtm_long <- bind_rows(gtm_list) %>% 
    filter(!is.na(Species))
unique(gtm_long$Species)
unique(gtm_long$QAQC)
# View(filter(gtm_long, QAQC == "<-3> (CSM)"))

gtm_qaqc <- gtm_long %>% 
    filter(QAQC != "<-3> (CSM)")
```

```{r}
gtm_cover <- gtm_qaqc %>% 
    select(all_of(grouping_cols2), Species, Cover) %>% 
    pivot_wider(names_from = Species,
                values_from = Cover)

gtm_density <- gtm_qaqc %>% 
    select(all_of(grouping_cols2), Species, Density) %>% 
    filter(!is.na(Density)) %>% 
    mutate(Species = paste0("Density_", Species)) %>% 
    pivot_wider(names_from = Species,
                values_from = Density)

gtm_height <- gtm_qaqc %>% 
    select(all_of(grouping_cols2), Species, 'Canopy Height') %>% 
    filter(!is.na('Canopy Height')) %>%
    mutate(Species = paste0("Canopy Height_", Species)) %>% 
    pivot_wider(names_from = Species,
                values_from = 'Canopy Height')

gtm_other_id <- gtm_qaqc %>% 
    select(all_of(grouping_cols2), all_of(other_id_cols)) %>% 
    distinct()

gtm_all <- full_join(gtm_other_id, gtm_cover, 
                     by = all_of(grouping_cols2)) %>% 
    full_join(., gtm_density, 
                     by = all_of(grouping_cols2)) %>% 
    full_join(., gtm_height, 
                     by = all_of(grouping_cols2))

write.csv(gtm_all, here::here("data", "gtm_namaste_template.csv"),
          row.names = FALSE,
          na = "")
```


```{r}
save(niw_all, niw_qaqc, niw_cover, niw_density, niw_height,
     gtm_all, gtm_qaqc, gtm_cover, gtm_density, gtm_height,
     file = here::here("data", "NIW_GTM.RDS"))
```

