---
title: "Coordinate Mapping"
author: "Kim Cressman"
date: "2022-04-28; latest update `r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_depth: 2
        code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

Really messy script to read in a single file for each reserve and pull GPS coordinates for their stations. While I'm at it, I'll see how many have multiple values for the same station (like GND's - eek). But I'll average coordinates together for this mapping purpose.  

```{r}
library(tidyverse)
library(readxl)
library(leaflet)
library(rgdal)
library(stringi)  # for stri_sub_replace()
```

# Setup  

### Output prep  

```{r}
all_coords <- tibble(Reserve = character(),
                     SiteID = character(),
                     Lat = numeric(),
                     Long = numeric())
```


### Helpful functions  

Because I'll be doing this repetitively on similar data files....  

```{r}
# starting path for data
path_start <- here::here("submitted_data", "data")

source(here::here("R", "sourced", "00_helper_functions.R"))
```



# Reserve-by-reserve  

## CDMO-formatted data  

### APA  

```{r}
file_in <- here::here(path_start, "APA", "APAVEG2020.xlsx")
apa <- read_cdmo(file_in)
bind_coords(apa)
```

### CBM - Jug Bay  

There seems to be a longitude typo at RR-T2-P5; first two digits past decimal are .77 but for other plots are .71. In first maps, using these coords (.77), P5 was quite far away from the rest of the transect. Fixing in code.  

-76.77092 --> -76.71092

```{r}
file_in <- here::here(path_start, "CBM", "Veg", "Jug Bay",
                      "CDMO EV Data.xlsx")
cbm_jb <- read_cdmo(file_in, worksheet = "2020") %>% 
    select(Reserve, SiteID, TransectID, PlotID, Lat, Long) %>% 
    distinct() %>% 
    mutate(Long = -1 * Long,
           Long = case_when(Long == -76.77092 ~ -76.71092,
                            TRUE ~ Long)) 

bind_coords(cbm_jb)
```



### CBV  

originally had to add "Other" as a column name in the file because there was text lower down in the column and this caused problems later. On 5/25 I updated the read_cdmo function to accomodate and it seems to work fine now.      

```{r}
file_in <- here::here(path_start, "CBV", "CBNERRVA VEG Data GI Reserve for Chris Peter.xlsx")
cbv <- read_cdmo(file_in) 
bind_coords(cbv)
```

### JAC  

```{r}
file_in <- here::here(path_start, "JAC", "2021", "JAC 2021 Marsh Vegetation.xlsx")
jac <- read_cdmo(file_in)
bind_coords(jac)
```


### NIW  

In this file, 'Lat' is repeated in the 'Long' column for Segment B, Plot 3-3. 

Per email with Tracy on 5/16/22, coordinates for this plot are:  
33.34045339	-79.2030272  


```{r}
file_in <- here::here(path_start, "NIW", "NIWVEG2020.xlsx")
niw <- read_cdmo(file_in) %>% 
    mutate(id = paste(SiteID, TransectID, PlotID, sep = "-"),
           Lat = case_when(id == "Segment B-3-3" ~ 33.34045339,
                           TRUE ~ Lat),
           Long = case_when(id == "Segment B-3-3" ~ -79.2030272,
                            TRUE ~ Long))


bind_coords(niw)
```


### GTM  

```{r}
file_in <- here::here(path_start, "GTM", "GTMVEG2020.xlsx")
gtm <- read_cdmo(file_in)
bind_coords(gtm)
```

### NOC  

Separate folders for different reserve components.  

Transects are parallel to shore rather than perpendicular.  

HAS NAs for LAT AND LONG which I was using above because other reserves didn't all have a number fo transect, plot, and/or cover. I don't know how else to make R recognize the number of header rows so I may have to just write a modified function for reserves like this.  

RC component has some leading and trailing spaces in Lat (and maybe Long) - removed spaces through the bind_coords function but check to see if it's a problem in other columns. This prevented the cells from being converted to numeric.  

```{r}
file_in <- here::here(path_start, "NOC", "Masonboro_Island_component",
                      "NOC_MI_2020.xlsx")
noc_mi <- read_cdmo(file_in, skip = 5)  # this is correct


file_in <- here::here(path_start, "NOC", "Rachel_Carson_component",
                      "NOC_RC_2019.xlsx")
noc_rc <- read_cdmo(file_in, skip = 5) # this is correct


file_in <- here::here(path_start, "NOC", "Zekes_Island_component",
                      "NOC_ZI_2019.xlsx")
noc_zi <- read_cdmo(file_in, skip = 5) # also worked correctly

bind_coords(noc_mi)
bind_coords(noc_rc)
bind_coords(noc_zi)
```


### GND  

```{r}
file_in <- here::here(path_start, "GND", "Veg Data and Metadata",
                      "GNDVEG2020.xlsx")
gnd <- read_cdmo(file_in)
bind_coords(gnd)
```


### MAR  

Somethings weird with these coordinates; they're not making transects when all coords for a plot are averaged together.  

```{r}
file_in <- here::here(path_start, "MAR", "Vegetation and Metadata",
                      "MARVEG2020_marsh_04.22.2021.xlsx")
mar <- read_cdmo(file_in)
```

```{r}
mar_plots <- mar %>% 
    select(SiteID, TransectID, PlotID) %>% 
    distinct()
```

MAR appears to have 325 distinct plots.  # transects per site:  

```{r}
mar_plots %>% 
    group_by(SiteID) %>% 
    summarize(n_transects = length(unique(TransectID))) %>% 
    knitr::kable()
```

5 sites with 5 transects each = 25 transects total.  

Number plots per transect:  

```{r}
mar_plots %>% 
    group_by(SiteID, TransectID) %>% 
    summarize(n_plots = length(unique(PlotID))) %>% 
    knitr::kable()
```

Most sites have 11 plots per transect; at HF2 there are 21 plots per transect.  


```{r}
mar %>% 
    select(SiteID, TransectID, PlotID, Date) %>% 
    group_by(SiteID, TransectID, PlotID) %>% 
    summarize(n_readings = length(unique(Date))) %>% 
    arrange(n_readings) %>% 
    knitr::kable()
```

```{r}
# leaflet(all_coords) %>% 
#     addProviderTiles(providers$Esri.WorldImagery) %>% 
#     addCircleMarkers(lat = ~Lat, lng = ~Long,
#                      weight = 2,
#                      radius = 2,
#                color = "orange",
#                popup = ~paste(Reserve, SiteID, TransectID, PlotID,
#                               sep = "-")) %>% 
#     addScaleBar()
```

Based on the interview with the reserve, all appears to be correct. The transects are 20m long and 2m apart, so plots are close together and with bigger points on the map, they couldn't be delineated. With smaller points, the transects become apparent.  

```{r}
bind_coords(mar)
```

### ELK  

```{r}
file_in <- here::here(path_start, "ELK", "ELKVEG2016.FINAL.xlsx")
elk <- read_cdmo(file_in)
bind_coords(elk)
```

### KAC  

originally had to add a 0 in the top QAQC column because otherwise it didn't get read in and caused problems. updated `read_cdmo` function on 5/25/2022 and problem is now fixed.  

```{r}
file_in <- here::here(path_start, "KAC", "Recent data",
                      "KACVEG2021.xlsx")
kac <- read_cdmo(file_in)
bind_coords(kac)
```


### SOS  

CDMO format, but different years in different worksheets. Will have to specify worksheet.  

Wait, they don't have lat/longs in this file! Must be in metadata somewhere. Oh, there's some in the 'Elevations' file. Not sure if PlotIDs are the same; will verify later but will work with those for now.   

```{r}
file_in <- here::here(path_start, "SOS", 
                      "SOS_Elevations_2010_2021.xlsx")

sos <- read_xlsx(file_in) %>% 
    mutate(Reserve = "SOS") %>% 
    select(Reserve, 
           SiteID = Site,
           TransectID,
           PlotID,
           Lat = LocalLatit,
           Long = LocalLongi)

bind_coords(sos)
```


### DEL  

CDMO format, but different reserve components in different worksheets. Will have to specify worksheet.  

```{r}
file_in <- here::here(path_start, "DEL", "Veg Data and Metadata",
                      "2018 DEL Veg Monitoring Raw Data.xlsx")
del_sjr <- read_cdmo(file_in, worksheet = "SJR") %>% 
    mutate(Reserve = "DEL-SJR")
del_bcr <- read_cdmo(file_in, worksheet = "BCR") %>% 
    mutate(Reserve = "DEL-BCR")
bind_coords(del_sjr)
bind_coords(del_bcr)
```


### LKS  

```{r}
file_in <- here::here(path_start, "LKS", "LKS NERR VegMonitoring 2014-2020_PROVISIONAL.xlsx")
lks <- read_cdmo(file_in, worksheet = "2020") %>% 
    rename(Lat = Latitude,
           Long = Longitude)
bind_coords(lks)
```



## Other long-format reserves  

### HUD - Tivoli  

Coords use UTM-18  

OTN 3B latitude is messed up (coming out as 23.99, which puts it in the Caribbean) - looks like the first digit was typed as a 2 and should have been a 4. Correcting here.    

```{r}
file_in <- here::here(path_start, "HUD", "HUD Tivoli Veg Data",
                      "VegMon Data 2020.xlsx")
prep <- read_xlsx(file_in, sheet = "Data Entry", n_max = 10)
names(prep)[3:4] <- c("northing", "easting")
hud_tiv <- read_xlsx(file_in, sheet = "Data Entry", skip = 2, col_names = FALSE)
names(hud_tiv) <- names(prep)

hud_tiv <- hud_tiv %>% 
    mutate(northing = case_when(PlotID == "OTN-3B" ~ 4654284,
                                TRUE ~ northing))
rm(prep)

points <- hud_tiv %>% 
    select(easting, northing)
# code below modified from stack overflow https://stackoverflow.com/a/30018607
sputm <- SpatialPoints(points, proj4string=CRS("+proj=utm +zone=18 +datum=WGS84")) 
spgeo <- spTransform(sputm, CRS("+proj=longlat +datum=WGS84"))
lnlt <- data.frame(coordinates(spgeo))

hud_tiv <- hud_tiv %>% 
    mutate(Lat = lnlt$northing,
           Long = lnlt$easting,
           Reserve = "HUD",
           SiteID = "TIV") %>% 
    separate(PlotID, into = c("TransectID", "PlotID"), sep = "-") %>% 
    select(Reserve, SiteID, TransectID, PlotID, Lat, Long, northing, easting)

bind_coords(hud_tiv)
```


### HUD - Piermont  

```{r}
file_in <- here::here(path_start, "HUD", "HUD Piermont Veg Data",
                      "Piermont VegMon Data 2020.xlsx")
prep <- read_xlsx(file_in, sheet = "Data Entry", n_max = 10)
names(prep)[5:6] <- c("northing", "easting")
hud_pmn <- read_xlsx(file_in, sheet = "Data Entry", skip = 2, col_names = FALSE)
names(hud_pmn) <- names(prep)
rm(prep)

hud_pmn <- hud_pmn %>% 
    filter(!is.na(northing))

points <- hud_pmn %>% 
    select(easting, northing)
# code below modified from stack overflow https://stackoverflow.com/a/30018607
sputm <- SpatialPoints(points, proj4string=CRS("+proj=utm +zone=18 +datum=WGS84")) 
spgeo <- spTransform(sputm, CRS("+proj=longlat +datum=WGS84"))
lnlt <- data.frame(coordinates(spgeo))

hud_pmn <- hud_pmn %>% 
    mutate(Lat = lnlt$northing,
           Long = lnlt$easting,
           Reserve = "HUD") %>% 
    separate(PlotID, into = c("TransectID", "SiteID", "PlotID"), sep = "-",
             extra = "merge") %>% 
    select(Reserve, SiteID, TransectID, PlotID, Lat, Long)

bind_coords(hud_pmn)
```


## NE project reserve packets  

GRB, NAR, WEL, WQB

### GRB  

Latitude and Longitude columns are blank; CP sent a CDMO formatted file    

```{r}
file_in <- here::here(path_start, "GRB", "GRBVEG2016.xlsx")
grb <- read_cdmo(file_in)
bind_coords(grb)
```

### NAR  

Latitude and Longitude columns are blank; using CDMO 2012 instead    

It seems that what's in the lat/long columns here was actually degrees-minutes-seconds format, entered into the decimal degree field. So I need to do some separating and converting.  

```{r}
file_in <- here::here(path_start, "NAR", "NARVEG2012", "NARVEG2012.xlsx")
nar <- read_cdmo(file_in)
```


```{r}
nar_coords <- nar %>% 
    select(Reserve, SiteID, TransectID, PlotID, Lat, Long) %>% 
    distinct() %>% 
    mutate_all(as.character) %>% 
    separate(Lat, into = c("Lat_deg", "Lat_min"), sep = "[.]") %>% 
    separate(Long, into = c("Long_deg", "Long_min"), sep = "[.]") %>% 
    mutate(Lat_min = stri_sub_replace(Lat_min, 3, 2, value = "."),
           Long_min = stri_sub_replace(Long_min, 3, 2, value = ".")) %>% 
    separate(Lat_min, into = c("Lat_min", "Lat_sec"), sep = "[.]") %>% 
    separate(Long_min, into = c("Long_min", "Long_sec"), sep = "[.]") %>% 
    mutate(Lat_sec = stri_sub_replace(Lat_sec, 3, 2, value = "."),
           Long_sec = stri_sub_replace(Long_sec, 3, 2, value = ".")) %>% 
    mutate(across(c(Lat_deg, Lat_min, Lat_sec, 
                    Long_deg, Long_min, Long_sec), 
                  as.numeric)) %>% 
    mutate(Lat = round(Lat_deg + (Lat_min/60) + (Lat_sec/3600), 5),
           Long = round(-1*((-1*Long_deg) + (Long_min/60) + (Long_sec/3600)), 5))

nar2 <- nar_coords %>% 
    mutate(across(c(TransectID, PlotID), as.numeric))

bind_coords(nar2)
```


### WEL  

lat/long columns are sparsely populated; using their 2020 CDMO-formatted file instead  

```{r}
file_in <- here::here(path_start, "WEL", "WEL Vegetation 2020 - CDMO Version.xlsx")
wel <- read_cdmo(file_in)
bind_coords(wel)
```

### WQB  

lat/long blank; using CDMO version instead  

lat/long were switched for S4. Correcting here with code; watch out for it when moving data to new template.  

```{r}
file_in <- here::here(path_start, "WQB", "WQBVEG2016", "WQBVEG2016.xlsx")
wqb <- read_cdmo(file_in)

wqb2 <- wqb %>% filter(Lat < 0) %>% 
    rename(Lat2 = Lat,
           Long2 = Long) %>% 
    rename(Lat = Long2,
           Long = Lat2)

wqb <- wqb %>% 
    filter(Lat > 0) %>% 
    bind_rows(wqb2)

bind_coords(wqb)
```


## Other wide-format reserves  
 

### ACE  

no lat/long columns in data file - is in Elevations file  

so being wide format doesn't matter  

```{r}
file_in <- here::here(path_start, "ACE", 
                      "Plot Elevations.xlsx")
ace <- read_xlsx(file_in) %>% 
    separate(ID, into = c("Reserve", "TransectID", "PlotID"), sep = "_") %>% 
    mutate(SiteID = "EI") %>% 
    select(-'Elevation (m)')

bind_coords(ace)
```


### CBM - Otter Point Creek  

```{r}
file_in <- here::here(path_start, "CBM", "Veg", "Otter Point Creek", 
                      "MASTER_OPC_EV_Transect_Data_060512_LC.xls")

# deal with headers and reading in
cbm_prep <- read_xls(file_in, sheet = "Site Descriptions",
                    skip = 10)
names(cbm_prep)[4:5] <- c("Lat", "Long")

cbm_opc <- read_xls(file_in, sheet = "Site Descriptions",
                     skip = 12,
                     col_names = FALSE)
names(cbm_opc) <- names(cbm_prep)
rm(cbm_prep)


# deal with data frame
cbm_opc <- cbm_opc %>% 
    fill(c(Transect, 'Current Transect Name')) %>% 
    mutate(Reserve = "CBM",
           SiteID = "OPC",
           Long = Long * -1) %>% 
    filter(!is.na(Plot)) %>% 
    select(Reserve, SiteID,
           TransectID = 'Current Transect Name',
           PlotID = Plot,
           Lat, Long)

bind_coords(cbm_opc)
```


### CBM - Monie Bay  

Appears to be old typo for MCMST3 P2 and P5. Per Kyle: I would be willing to bet if you changed say, P2 from 48.6 to 46.6 it would line up.  

```{r}
file_in <- here::here(path_start, "CBM", "Veg", "Monie Bay", 
                      "MASTER MBR_EV_Transect_Data_Current_2017.xlsx")


cbm_prep <- read_xlsx(file_in, sheet = "Site Description",
                    skip = 9)
names(cbm_prep)[3:4] <- c("Lat_dms", "Long_dms")

cbm_mb <- read_xlsx(file_in, sheet = "Site Description",
                     skip = 11,
                     col_names = FALSE)
names(cbm_mb) <- names(cbm_prep)
rm(cbm_prep)

cbm_mb <- cbm_mb %>% 
    fill(Transect) %>% 
    filter(!is.na(Plot)) %>% 
    mutate(Reserve = "CBM",
           SiteID = "MB") %>% 
    select(Reserve, SiteID,
           TransectID = Transect,
           PlotID = Plot,
           Lat_dms, Long_dms)

# convert gps coordinates
cbm_mb2 <- cbm_mb %>% 
    mutate(Lat_dms2 = str_replace_all(Lat_dms, "[^[:digit:][\\.]]+", " "),
           Lat_dms2 = str_trim(Lat_dms2),
           Long_dms2 = str_replace_all(Long_dms, "[^[:digit:][\\.]]+", " "),
           Long_dms2 = str_trim(Long_dms2),
           ID = paste(TransectID, PlotID, sep = "-")) %>% 
    separate(Lat_dms2, into = c("Lat_deg", "Lat_min", "Lat_sec"),
             sep = " ") %>% 
    separate(Long_dms2, into = c("Long_deg", "Long_min", "Long_sec"),
             sep = " ") %>% 
    mutate(across(Lat_deg:Long_sec,
                  as.numeric)) %>% 
    mutate(Lat_sec = case_when(ID == "MCMST3-P2" ~ 46.9,
                               ID == "MCMST3-P5" ~ 47.8,
                               TRUE ~ Lat_sec)) %>% 
    mutate(Lat = round(Lat_deg + (Lat_min/60) + (Lat_sec/3600), 5),
           Long = round(-1*(Long_deg + (Long_min/60) + (Long_sec/3600)), 5))  
```


```{r}
bind_coords(cbm_mb2)
```

Write out a csv of all coordinates:  

```{r}
write.csv(all_coords, here::here("data", "all_plot_locations.csv"),
          row.names = FALSE)
```


# Number plots per reserve   

See what we've got:  

```{r}
all_coords %>% 
    group_by(Reserve) %>% 
    tally() %>% 
    knitr::kable(align = "l")
```

# Map!  

```{r, fig.height = 7, fig.width = 10}
leaflet(all_coords) %>% 
    addProviderTiles(providers$Esri.WorldImagery) %>% 
    addCircles(lat = ~Lat, lng = ~Long,
               weight = 2,
               radius = 2,
               color = "orange",
               popup = ~paste(Reserve, SiteID, TransectID, PlotID,
                              sep = "-")) %>% 
    addScaleBar()
```

