---
title: "NMST file checks"
format: html
execute:
  echo: false
---
```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(glue)
```


UNLESS a warning message exists below the name of each file, the following checks have been performed successfully:   

-  sheet names are correct, and in the correct order  
-  all identifying column names exist in the Cover sheet (Reserve, SiteID, TransectID, PlotID, Year, Month, Day, Orthometric_Height, Height_Relative_to_MLLW, Distance)
-  there aren't any duplicate rows of date-site-transect-plot  
-  species names match between sheets: names in density and height sheets should be subset of those on cover sheet  

These have not yet been coded but will be:  
-  station table sheet contains all stations used in cover sheet  
-  no columns are named "NA" or "F_NA" (this happened occasionally when Reserves would have a blank for Species on dates that a plot was not read)  

```{r}
sht_names <- c("Station_Table", "Species_Names", "Cover", "Density", "Height")

id_cols <- c("Reserve", "SiteID", 
             "Year", "Month", "Day",
             "TransectID", "PlotID", 
             "Orthometric_Height", "Height_Relative_to_MLLW",
             "Notes")

stn_tbl_cols <- c("Reserve", "SiteID", 
                  "TransectID", "PlotID", 
                  "Lat", "Long", "Type",
                  "Orthometric_Height", "Height_Relative_to_MLLW",
                  "Distance", "SSAM-1")
```


## What are the files in the directory  

```{r}
path <- here::here("from_google_drive",
                        "Namaste-ish straight out of R")
fls <- list.files(path)
fls2 <- paste0(path, "/", fls)

print(fls)
```



## Run checks on each  

```{r}
for(i in seq_along(fls)){
    # print some dividers
    cat("\n----------------------- \n")
    cat("----------------------- \n")
    
    # source script of checks  
    try(
        source(here::here("R", "sourced", "nmst-file-check-helper.R"))
    )

}
```

