---
title: "CDMO-to-NMST"
author: "Kim Cressman"
date: "latest update: 6/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)
library(writexl)


# function to pick first non-NA reading for any given plot ("at establishment")
establ <- function(df, x){
    
    x_vec <- df %>% 
        select({{x}})
    
    if(sum(!is.na(x_vec)) == 0){
        out <- df %>% 
            select(SiteID, TransectID, PlotID, {{x}}) %>% 
            distinct() 
    } else {
        out <- df %>% 
            select(Date, SiteID, TransectID, PlotID, {{x}}) %>% 
            distinct() %>% 
            group_by(SiteID, TransectID, PlotID) %>% 
            filter(!is.na({{x}})) %>%
            filter(Date == min(Date)) %>% 
            select(-Date)
    }
        
    out
}
```

# Identify file to read in  

In a Shiny app, let someone pick this interactively. Here, hard-code. Using MAR because it's the first full working file I have. NOC-ZI is also good.  

```{r}
# reserve <- "APA"  # contains Average and Maximum Canopy Height
# reserve <- "CBM-JB"  # Maximum Canopy Height
# reserve <- "GND"  # completed 6/17/12
# reserve <- "KAC"  # completed 6/17/22
reserve <- "LKS"
# reserve <- "MAR"  # completed 6/17/22
# reserve <- "NOC-MI"
# reserve <- "NOC-RC"
# reserve <- "NOC-ZI"
```

```{r}
# path <- here::here("wrangled_data", "CDMO", "MAR_CDMO.xlsx")
path <- here::here("wrangled_data", "CDMO", paste0(reserve, "_CDMO.xlsx"))

# out_path <- here::here("wrangled_data", "NMST", "MAR_NMST.xlsx")
out_path <- here::here("wrangled_data", "NMST", paste0(reserve, "_NMST.xlsx"))
```


Renaming possibilities:  

```{r}
col_matching <- c("Orthometric_Height" = "Orthometric Height",
                  "Height_Relative_to_MLLW" = "Height Relative to MLLW")
```

Read in data and rename:  

```{r}
dat <- read_xlsx(path,
                 guess_max = 2500) %>% 
    rename(any_of(col_matching))
```

If QAQC column doesn't exist, add one full of NAs.  

```{r}
if(!exists("QAQC", dat)){
  dat$QAQC <- NA  
}
```


## Identifiers  

Should be the same between all sheets  

```{r}
id_names <- c("Reserve", "SiteID", 
             "Year", "Month", "Day",
             "TransectID", "PlotID",
             "Orthometric_Height", "Height_Relative_to_MLLW",
             "Distance")

# only use ID columns that exist in the data frame
id_names <- id_names[which(id_names %in% names(dat))]
```

## Deal with notes separately  

If there's not a Notes column, make one full of NAs.  

```{r}
if(exists("Notes", dat)){
    notes <- dat %>% 
        filter(!is.na(Notes))
    c_notes <- notes %>% 
        group_by(SiteID, Year, Month, Day, TransectID, PlotID) %>% 
        summarize(Notes = str_c(unique(Notes), collapse = "; "))
} else {
    c_notes <- dat %>% 
        group_by(SiteID, Year, Month, Day, TransectID, PlotID) %>% 
        summarize(Notes = NA_character_)
}
```

# Cover sheet  

NOTE this sheet also has average canopy height and density. Replicates are on the other sheets.  

## Cover data  

```{r}
covr <- dat %>% 
    select(all_of(id_names), Species, Cover, QAQC)

# make sure there aren't any dupes  
janitor::get_dupes(covr, -Cover)
```

```{r}
# pivot
covr_wide <- covr %>% 
    rename("F" = QAQC) %>% 
    pivot_wider(names_from = Species,
                values_from = c(Cover, "F"))

# get the 'Cover' prefix out of the names
names(covr_wide) <- str_replace(names(covr_wide), "^(Cover_)", "")
```

## Density data  

```{r}
dens <- dat %>% 
    select(all_of(id_names), Species, Density) %>% 
    filter(!is.na(Density)) %>% 
    pivot_wider(names_from = Species,
                values_from = Density,
                names_prefix = "Density_")
```


## Height data  

```{r}
hts <- dat %>% 
    select(all_of(id_names), Species, Height = "Canopy Height") %>% 
    filter(!is.na(Height)) %>% 
    pivot_wider(names_from = Species,
                values_from = Height,
                names_prefix = "Average Canopy Height_")
```


## Mangrove height and diameter  

if applicable.  

```{r}
spp_all <- sort(unique(dat$Species))

# are mangroves present in the species list
mangroves <- sum(str_detect(spp_all, "Avic")) + sum(str_detect(spp_all, "Lagunc")) + sum(str_detect(spp_all, "Rhizo"))

# what are those species
mang_spp <- spp_all[str_which(spp_all, "Avic")]

if(mangroves > 0){
    dat_mang_ht <- dat %>% 
        filter(Species %in% mang_spp) %>% 
        select(all_of(id_names), Species, Height) %>% 
        pivot_wider(names_from = Species,
                values_from = Height,  # Height in original dataset is mangrove height
                names_prefix = "Height_")
    
    dat_mang_diam <- dat %>% 
        filter(Species %in% mang_spp) %>% 
        select(all_of(id_names), Species, Diameter) %>% 
        pivot_wider(names_from = Species,
                values_from = Diameter,  # Height in original dataset is mangrove height
                names_prefix = "Diameter_")
    
    dat_mang <- full_join(dat_mang_ht, dat_mang_diam)
}
```


## Join them all  

```{r}
covr_sheet <- full_join(covr_wide, dens) %>% 
    full_join(hts) %>% 
    full_join(c_notes) %>% 
    select(all_of(id_names), Notes, everything()) %>% 
    relocate(Notes, .after = Distance) %>% 
    relocate(starts_with("F_"), .after = last_col())

# join in mangrove heights and diameters too
if(mangroves > 0){
    covr_sheet <- full_join(covr_sheet, dat_mang) %>% 
        relocate(starts_with("F_"), .after = last_col())
}
```


# Density sheet  

Subset from covr_sheet to ensure equal number of rows between sheets  

```{r}
dens_sheet <- covr_sheet %>% 
    select(all_of(id_names), Notes,
           starts_with("Density_"))

# get the 'Density' prefix out of the names
names(dens_sheet) <- str_replace(names(dens_sheet), "^(Density_)", "")
```


# Height sheet  

Subset from covr_sheet to ensure equal number of rows between sheets  

```{r}
hts_sheet <- covr_sheet %>% 
    select(all_of(id_names), Notes,
           starts_with("Average Canopy Height_"))

# get the 'Density' prefix out of the names
names(hts_sheet) <- str_replace(names(hts_sheet), "^(Average Canopy Height_)", "")
```


# Station Table  

Everything in this table is the first recorded value.  


```{r}
# there must be a better way to do this, but it works.

stns <- dat %>% 
    select(Reserve, SiteID, TransectID, PlotID) %>% 
    distinct()
lats <- establ(dat, Lat)
longs <- establ(dat, Long)
ssam1 <- establ(dat, "SSAM-1")
dists <- establ(dat, Distance)
orth_hts <- establ(dat, Orthometric_Height) 
mllw_hts <- establ(dat, Height_Relative_to_MLLW)
types <- establ(dat, Type)

stn_tbl <- full_join(stns, lats) %>% 
    full_join(longs) %>% 
    full_join(types) %>% 
    full_join(orth_hts) %>% 
    full_join(mllw_hts) %>%
    full_join(dists) %>% 
    full_join(ssam1)

```



# Species List  

```{r}
spp_names <- dat %>% 
    select(Species, Type) %>% 
    distinct() %>% 
    arrange(Species)
```




# Write to Excel  

```{r}
write_xlsx(list(Station_Table = stn_tbl,
                Species_Names = spp_names,
                Cover = covr_sheet,
                Density = dens_sheet,
                Height = hts_sheet),
           path = out_path,
           format_headers = TRUE)
```

